---
title: "PF_Giovanni preliminary analyses"
author: 'Volcano plots - Metabolomics'
date: "SJ"
output:
  html_document:
    toc: yes
    number_sections: true
    toc_depth: 5
  pdf_document:
    toc: yes
    number_sections: true
    toc_depth: 5
urlcolor: blue
linkcolor: red
---
```{r global_options, include=FALSE}
rm(list = ls())
# if (!requireNamespace("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# BiocManager::install("EnhancedVolcano")
library(EnhancedVolcano)
library(dplyr)
library(forecast)
library(VennDiagram)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, warning=FALSE,
                      fig.width = 6, fig.asp = 0.8, out.width = "80%")
#knitr::opts_knit$set(root.dir = "G:/My Drive/Prolonged_Fasting_Study")
knitr::opts_knit$set(root.dir = "../..") # PFS directory
theme_update(plot.title = element_text(hjust = 0.5)) # center plots
```


# Useful functions
```{r, message=FALSE}
#```{r results='hide', fig.keep='high'}
source("./SCRIPTS/Giovanni/Metabolomics functions_SJ_GF.R")
```

# Working directory and datasets
```{r}
setwd("./DATA")
df_metabolomics=read.csv("PF_Metabolomics_SJedit_added_10_unknowns.csv",header=T)
```

# Format dataset structure

```{r}
# Get named group factor
group<-factor(sub(".*\\.", "", colnames(df_metabolomics)[-1]))
names(group)<-colnames(df_metabolomics)[-1]

#Rownames as first column
rownames(df_metabolomics)<-df_metabolomics[,1]
df_metabolomics<-df_metabolomics[,-1]
```

# Differential Expression results

```{r}
### run analyses
results_BL_W1 = create_results_table(df=df_metabolomics,group1="W1",group2="BL") # BL vs W1
results_BL_EF = create_results_table(df=df_metabolomics,group1="EF",group2="BL") # BL vs EF
results_EF_ER = create_results_table(df=df_metabolomics,group1="ER",group2="EF") # EF vs ER
results_BL_ER = create_results_table(df=df_metabolomics,group1="ER",group2="BL") # BL vs ER
```

# Volcano Plots

## BL to W1

```{r}
# Adjusted
EnhancedVolcano(results_BL_W1, pCutoff = 0.05,FCcutoff = 1,
                lab = rownames(results_BL_EF), labSize = 2,
                x = 'log2FC', y = 'FDR', title="BLW1 Volcano Plot",
                subtitle="Using FDR (adjusted) p-values",
                selectLab = rownames(results_BL_W1)[which(abs(results_BL_W1$log2FC)>=1 | results_BL_W1$FDR<=0.05)],
                legendPosition = 'right',legendLabSize = 9,legendIconSize = 3,
                boxedLabels = TRUE, labFace='bold',
                drawConnectors = TRUE,
                widthConnectors = 0.75)
```

## BL to EF

```{r}
# Unadjusted
EnhancedVolcano(results_BL_EF, pCutoff = 1e-03,FCcutoff = 1,
                lab = rownames(results_BL_EF), labSize = 2,
                x = 'log2FC', y = 'pval', title="BLEF Volcano Plot",
                subtitle="Using unadjusted p-values",
                selectLab = rownames(results_BL_EF)[which(abs(results_BL_EF$log2FC)>=1 | results_BL_EF$pval<=1e-03)],
                legendPosition = 'right',legendLabSize = 9,legendIconSize = 3,
                boxedLabels = TRUE, labFace='bold',
                drawConnectors = TRUE,
                widthConnectors = 0.75)

# Adjusted
EnhancedVolcano(results_BL_EF, pCutoff = 0.05,FCcutoff = 1,
                lab = rownames(results_BL_EF), labSize = 2,
                x = 'log2FC', y = 'FDR', title="BLEF Volcano Plot",
                subtitle="Using FDR (adjusted) p-values",
                selectLab = rownames(results_BL_EF)[which(abs(results_BL_EF$log2FC)>=1 | results_BL_EF$FDR<=0.05)],
                legendPosition = 'right',legendLabSize = 9,legendIconSize = 3,
                boxedLabels = TRUE, labFace='bold',
                drawConnectors = TRUE,
                widthConnectors = 0.75)

```

## EF to ER

```{r}
# Unadjusted
EnhancedVolcano(results_EF_ER, pCutoff = 1e-03,FCcutoff = 1,
                lab = rownames(results_EF_ER), labSize = 2,
                x = 'log2FC', y = 'pval', title="EFER Volcano Plot",
                subtitle="Using unadjusted p-values",
                selectLab = rownames(results_EF_ER)[which(abs(results_EF_ER$log2FC)>=1 | results_EF_ER$pval<=1e-03)],
                legendPosition = 'right',legendLabSize = 9,legendIconSize = 3,
                boxedLabels = TRUE, labFace='bold',
                drawConnectors = TRUE,
                widthConnectors = 0.75)

# Adjusted
EnhancedVolcano(results_EF_ER, pCutoff = 0.05,FCcutoff = 1,
                lab = rownames(results_EF_ER), labSize = 2,
                x = 'log2FC', y = 'FDR', title="EFER Volcano Plot",
                subtitle="Using FDR (adjusted) p-values",
                selectLab = rownames(results_EF_ER)[which(abs(results_EF_ER$log2FC)>=1 | results_EF_ER$FDR<=0.05)],
                legendPosition = 'right',legendLabSize = 9,legendIconSize = 3,
                boxedLabels = TRUE, labFace='bold',
                drawConnectors = TRUE,
                widthConnectors = 0.75)

```

## BL to ER

```{r}
# Unadjusted
EnhancedVolcano(results_BL_ER, pCutoff = 1e-03,FCcutoff = 1,
                lab = rownames(results_BL_ER), labSize = 2,
                x = 'log2FC', y = 'pval', title="BLER Volcano Plot",
                subtitle="Using unadjusted p-values",
                selectLab = rownames(results_BL_ER)[which(abs(results_BL_ER$log2FC)>=1 | results_BL_ER$pval<=1e-03)],
                legendPosition = 'right',legendLabSize = 9,legendIconSize = 3,
                boxedLabels = TRUE, labFace='bold',
                drawConnectors = TRUE,
                widthConnectors = 0.75)

# Adjusted
EnhancedVolcano(results_BL_ER, pCutoff = 0.05,FCcutoff = 1,
                lab = rownames(results_BL_ER), labSize = 2,
                x = 'log2FC', y = 'FDR', title="BLER Volcano Plot",
                subtitle="Using FDR (adjusted) p-values",
                selectLab = rownames(results_BL_ER)[which(abs(results_BL_ER$log2FC)>=1 | results_BL_ER$FDR<=0.05)],
                legendPosition = 'right',legendLabSize = 9,legendIconSize = 3,
                boxedLabels = TRUE, labFace='bold',
                drawConnectors = TRUE,
                widthConnectors = 0.75)
```

# Venn Diagrams

## Significant Fold-change and FDR p-val

```{r}
# Both conditions - FC and pval
fc_blw1 = rownames(results_BL_W1[(abs(results_BL_W1$log2FC)>=1) & 
                                       (abs(results_BL_W1$FDR)<=0.05),])
fc_blef = rownames(results_BL_EF[(abs(results_BL_EF$log2FC)>=1) & 
                                       (abs(results_BL_EF$FDR)<=0.05),])
fc_efer = rownames(results_BL_EF[(abs(results_EF_ER$log2FC)>=1) & 
                                       (abs(results_EF_ER$FDR)<=0.05),])
fc_bler = rownames(results_BL_ER[(abs(results_BL_ER$log2FC)>=1) & 
                                       (abs(results_BL_ER$FDR)<=0.05),])

setwd("./SCRIPTS/Giovanni/VennDiagrams")
venn.diagram(main='Delta changes Metabolites_FC and FDR',
  x = list(fc_blw1,fc_blef, fc_efer, fc_bler),
  category.names = c("BLW1","BLEF" , "EFER " , "BLER"),
  filename = 'VennD_Delta changes Metabolites.png',
  output=TRUE
)

blef_blw1_intrsct<-intersect(fc_blef, fc_blw1)
blef_efer_intrsct<-intersect(fc_blef, fc_efer)
```

## Significant Fold-change (only)

```{r}
# Only condition - FC
blef_sig_fc_pval = rownames(results_BL_EF[(abs(results_BL_EF$log2FC)>=1) & 
                                       (abs(results_BL_EF$FDR)>0.05),])
efer_sig_fc_pval = rownames(results_BL_EF[(abs(results_EF_ER$log2FC)>=1) & 
                                       (abs(results_EF_ER$FDR)>0.05),])
bler_sig_fc_pval = rownames(results_BL_ER[(abs(results_BL_ER$log2FC)>=1) & 
                                       (abs(results_BL_ER$FDR)>0.05),])

setwd("./SCRIPTS/Giovanni/VennDiagrams")
venn.diagram(main='Delta changes Metabolites_OnlyFC',
  x = list(blef_sig_fc_pval, efer_sig_fc_pval, bler_sig_fc_pval),
  category.names = c("BLEF" , "EFER " , "BLER"),
  filename = 'VennD_Delta changes Metabolites_OnlyFC.png',
  output=TRUE
)

intersect(blef_sig_fc_pval, efer_sig_fc_pval)
```

## Significant FDR p-val (only)

```{r}
# Only condition - FDR
blef_sig_fc_pval = rownames(results_BL_EF[(abs(results_BL_EF$log2FC)<1) & 
                                       (abs(results_BL_EF$FDR)<=0.05),])
efer_sig_fc_pval = rownames(results_BL_EF[(abs(results_EF_ER$log2FC)<1) & 
                                       (abs(results_EF_ER$FDR)<=0.05),])
bler_sig_fc_pval = rownames(results_BL_ER[(abs(results_BL_ER$log2FC)<1) & 
                                       (abs(results_BL_ER$FDR)<=0.05),])

setwd("./SCRIPTS/Giovanni/VennDiagrams")
venn.diagram(main='Delta changes Metabolites_OnlyFDR',
  x = list(blef_sig_fc_pval, efer_sig_fc_pval, bler_sig_fc_pval),
  category.names = c("BLEF" , "EFER " , "BLER"),
  filename = 'VennD_Delta changes Metabolites_OnlyFDR.png',
  output=TRUE
)

intersect(blef_sig_fc_pval, efer_sig_fc_pval)
```

# Save Significant Metabolites in RDA

```{r}
# Saved in Giovanni folder since we are dealing with metabolomics
setwd("./SCRIPTS/Giovanni")
save(fc_blw1, fc_blef,fc_efer,fc_bler,
     blef_blw1_intrsct, blef_efer_intrsct,
     file="RDA_Metabolites_FC.rda")
```