---
title: "PF_Clustering_without electryoltes etc"
author: ''
date: "SJ, AM"
output:
  html_document:
    toc: yes
    number_sections: true
    toc_depth: 5
  pdf_document:
    toc: yes
    number_sections: true
    toc_depth: 5
urlcolor: blue
linkcolor: red
---

```{r}
rm(list = ls()) ##AM reset workspace - good practice to run clean
library(ggplot2)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, warning=FALSE)
theme_update(plot.title = element_text(hjust = 0.5)) # center plots
```

# Working Directory & Datasets
```{r message=FALSE}

##AM please suppress trash outputs with chunk option: message=FALSE
##AM sometimes you need this to suppress stubborn outputs
#```{r results='hide', fig.keep='high'}

setwd("C:/Users/SJ/Desktop/Backup Mac/Desktop/STUDY/Andrius Prolonged Fasting")
fast_wide = read.csv("Blood_changes_across_timepoints_rows_wide_AM.csv")
fast_frac_wide = read.csv("Blood_changes_across_timepoints_rows_fractions_AM.csv")

# fast_wide = read.csv("../DATA/Blood_changes_across_timepoints_rows_wide_AM.csv",header=T,sep=",")
# 
# fast_frac_wide = read.csv("../DATA/Blood_changes_across_timepoints_rows_wide_fractions_AM.csv",header=T,sep=",")

# drop rows containing W1 timepoint - we will not use it
# fast_wide <- subset(fast_wide, Timepoint != "W1")
# fast_frac_wide <- subset(fast_frac_wide, Timepoint != "W1.f")

# working with fast_wide data
attach(fast_wide)
fast_wide$ID = as.factor(ID)
fast_wide$Timepoint=as.factor(Timepoint)

```

## Missing values?

```{r}

##AM add print() arguments to label each output
total_missing=sum(is.na(fast_wide))
(NAs_observations=which(rowMeans(is.na(fast_wide))!=0))
rowMeans(is.na(fast_wide[NAs_observations,])) # out of 50 observations
rowSums(is.na(fast_wide[NAs_observations,]))
fast_wide[NAs_observations,"Timepoint"] # All NA-heavy ones W1 so it's fine.
# 3 BL (1st, 21st and 37th), 1 ER (40th obs) missing
```

# Only Baseline
```{r}
library(dplyr) #for data manipulation
# only baseline (for clustering participants based on similarity)
bl_data=fast_wide[Timepoint=="BL",]

# Remove unwanted variables from clustering
baseline_data=select(bl_data,-matches("Weight|Waist|SBP|DBP|^Hb$|Ht|Neutr|Lymp|Mono|Sodium|Potassium|Chloride|Calcium|ALT|Triglycerides|^HDL$"))
sum(is.na(baseline_data)) # 3 missing values
(NAs_cols=which(colMeans(is.na(baseline_data))!=0)) # Bililrubin and UrineSG
# baseline=na.omit(baseline_data) # better removing Bilirubin and UrineSG instead of observations?

```

## Version 1: Without Bilirubin and UrineSG (missing)
```{r}
## Version 1: Without Bilirubin, UrineSG
baseline_1 = baseline_data[,-NAs_cols]
```

### Scaling
```{r}
### TRIAL : CLUSTERING

# https://uc-r.github.io/kmeans_clustering
library(cluster)    # clustering algorithms
library(factoextra) # clustering algorithms & visualization

scaled.baseline_1= scale(baseline_1[,-c(1,2)]) #removing timepoint and ID cols.
rownames(scaled.baseline_1)=baseline_1$ID
baseline_1_cor_mat=cor(scaled.baseline_1)
# View(baseline_1_cor_mat) #Uncomment to get table view in R
```

### Dissimilarity Matrix
```{r}
dist.mat = get_dist(scaled.baseline_1) # Euclidean default

### Dissimilarity matrix
fviz_dist(dist.mat, gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07"))
```

### Optimal number of clusters
```{r}

### Optimal number of clusters?
fviz_nbclust(scaled.baseline_1, kmeans, method = "wss") # 2 or 4?
fviz_nbclust(scaled.baseline_1, kmeans, method = "silhouette") # 4

# Tibshirani, Walter, Hastie
# gap_stat = clusGap(scaled.baseline_1, FUN = kmeans, K.max = 10, B = 50)
# fviz_gap_stat(gap_stat)
```

#### K-Means
```{r}

##AM let's plot several k-means to see visually
# K-means clustering, 2 clusters
baseline_1_kmeans2 = kmeans(scaled.baseline_1, 2,nstart=25)
fviz_cluster(baseline_1_kmeans2, data = scaled.baseline_1)

# K-means clustering, 3 clusters
baseline_1_kmeans3 = kmeans(scaled.baseline_1, 3,nstart=25)
fviz_cluster(baseline_1_kmeans3, data = scaled.baseline_1)

# K-means clustering, 4 clusters - last, to keep clusters
baseline_1_kmeans4 = kmeans(scaled.baseline_1, 4,nstart=25)
fviz_cluster(baseline_1_kmeans4, data = scaled.baseline_1)

```

#### Hierchical Clustering
```{r}
### Hierarchical clustering
baseline_1_hclust1 = hclust(dist.mat) #complete linkage
baseline_1_hclust2 = hclust(dist.mat,method='average')
baseline_1_hclust3 = hclust(dist.mat,method='single')
baseline_1_hclust4 = hclust(dist.mat, method = "ward.D2")

# Cut tree into 4 groups
grp1 = cutree(baseline_1_hclust1, k = 4)
grp2 = cutree(baseline_1_hclust2, k = 4)
grp3 = cutree(baseline_1_hclust3, k = 4)
grp4 = cutree(baseline_1_hclust4, k = 4)

##AM add ID column as a variable for handling after, column names
cluster_labels_df = data.frame(baseline_1$ID, baseline_1_kmeans2$cluster,
                               baseline_1_kmeans3$cluster,
                               baseline_1_kmeans4$cluster, 
                               grp1,grp2,grp3,grp4)
colnames(cluster_labels_df)=c("ID", "kmeans2", "kmeans3","kmeans4",
                              "hclust_comp", "hclust_avg", 
                              "hclust_sngl", "hclust_ward")

##SJ read.csv(filename,row.names=T,header=T) 
#SJ for importing written csv in next line

# write.csv(cluster_labels_df,"../DATA/Cluster labels baseline_without_e_AM.csv", row.names = F)
write.csv(cluster_labels_df,"C:/Users/SJ/Desktop/Backup Mac/Desktop/STUDY/Andrius Prolonged Fasting/Cluster labels baseline (4 clusters).csv", row.names = T)


##AM for some reason csv does not get save when I render html,
# only when I run this chunk in RStudio

# Visualize
# par(mfrow=c(2,2),pty='m') # Uncomment to see on one plot to compare
plot(baseline_1_hclust1, cex = 0.6) # plot tree complete linkage
rect.hclust(baseline_1_hclust1, k = 4, border = 2:5) # add rectangle

plot(baseline_1_hclust2, cex = 0.6) #average
rect.hclust(baseline_1_hclust2, k = 4, border = 2:5) 

plot(baseline_1_hclust3, cex = 0.6) #single
rect.hclust(baseline_1_hclust3, k = 4, border = 2:5) 

plot(baseline_1_hclust4, cex = 0.6) # Ward's method
rect.hclust(baseline_1_hclust4, k = 4, border = 2:5)
```

#### Stability of clusters
```{r results='hide', fig.keep='high'}

# Bootstrap to see stability of clusters
library(pvclust)
##AM I am using this to see BP values between rows (patients)
##AM complete and ward methods seem to be best with this data
baseline_1_pvclust1 = pvclust(scaled.baseline_1,
                              method.hclust="complete")
baseline_1_pvclust2 = pvclust(scaled.baseline_1,
                              method.hclust="ward.D2") 

baseline_1_pvclust3 = pvclust(t(scaled.baseline_1),
                              method.hclust="complete")
baseline_1_pvclust4 = pvclust(t(scaled.baseline_1),
                              method.hclust="ward.D2")
# AU around 95% implies strongly supported by data
plot(baseline_1_pvclust1)
pvrect(baseline_1_pvclust1, alpha = 0.90)
plot(baseline_1_pvclust2)
pvrect(baseline_1_pvclust2, alpha = 0.90)
plot(baseline_1_pvclust3)
pvrect(baseline_1_pvclust3, alpha = 0.90)
plot(baseline_1_pvclust4)
pvrect(baseline_1_pvclust4, alpha = 0.90)

```

#### Most appropriate clustering method?
```{r}
# Choice of appropriate clustering algorithm (G Brock 2008)
##AM only showing optimal scores and plots; hc and km come close, 2,3,4 clusters
library(clValid)
intern <- clValid(scaled.baseline_1, nClust = 2:11,
                  clMethods = c("hierarchical","kmeans","pam"), 
                  validation = "internal")
optimalScores(intern)
par(mfrow=c(1,3))
plot(intern)

stab <- clValid(scaled.baseline_1, nClust = 2:11,
                  clMethods = c("hierarchical","kmeans","pam"), 
                  validation = "stability")
optimalScores(stab)
# summary(stab)
par(mfrow=c(2,2))
plot(stab)

```

### PCA
```{r results='hide', fig.keep='high'}
##AM I would leave BMI as mprtant characteristic at BL and W1 - agreed.
res.pca <- prcomp(scaled.baseline_1)
# Visualize eigenvalues/variances
fviz_screeplot(res.pca, addlabels = TRUE, ylim = c(0, 50))
# Extract the results for variables
var = get_pca_var(res.pca)
# Contributions of variables to PC1
fviz_contrib(res.pca, choice = "var", axes = 1, top = 25)
# Contributions of variables to PC2
fviz_contrib(res.pca, choice = "var", axes = 2, top = 25)
# Control variable colors using their contributions to the principle axis
fviz_pca_var(res.pca, col.var="contrib",gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE # Avoid text overlapping
             ) + theme_minimal() + ggtitle("Variables - PCA")

##SJ To see interactions with only biomarkers - 
##SJ useful for model selection+resids diagnostics for gam/lm outputs with BMI as response.
### PCA without BMI and weight (only biomarkers)
res.pca <- prcomp(scaled.baseline_1[,-1])
# Visualize eigenvalues/variances
fviz_screeplot(res.pca, addlabels = TRUE, ylim = c(0, 50))
# Extract the results for variables
var = get_pca_var(res.pca)
# Contributions of variables to PC1
fviz_contrib(res.pca, choice = "var", axes = 1, top = 25)
# Contributions of variables to PC2
fviz_contrib(res.pca, choice = "var", axes = 2, top = 25)
# Control variable colors using their contributions to the principle axis
fviz_pca_var(res.pca, col.var="contrib",gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE # Avoid text overlapping
             ) + theme_minimal() + ggtitle("Variables - PCA")
```

```{r}

# Temporary break in code for testing
print("Early termination of code...")
knitr::knit_exit()

```

*** 

# BL $\to$ EF
```{r}
# BL to EF fractionalised-percentage changes 
# (Note: think about logs for skews ???)
blef_change=1+((fast_wide[Timepoint=="EF",-c(1,2)]-fast_wide[Timepoint=="BL",-c(1,2)])/fast_wide[Timepoint=="BL",-c(1,2)])
blef_change_data=select(blef_change,-matches("Weight|Waist|SBP|DBP|^Hb$|Ht|Neutr|Sodium|Potassium|Chloride|Calcium|ALT|Triglycerides|^HDL$"))
rownames(blef_change_data)=fast_wide[rownames(blef_change_data),"ID"]
# View(blef_change_data) #Uncomment to see table

# hist(blef_change_data[,"Weight"]) #Uncomment to see for skews/log-changes

(NAs_cols=which(colMeans(is.na(blef_change_data))!=0)) # Bililrubin and UrineSG similar to baseline_data
colSums(is.na(blef_change_data[,NAs_cols]))
```

## Version 1: Without Bilirubin and UrineSG (missing)
```{r}
## Version 1: Without Bilirubin, UrineSG
blef_1 = blef_change_data[,-NAs_cols]
```

### Scaling
```{r}
### TRIAL : CLUSTERING

scaled.blef_1= scale(blef_1)
rownames(scaled.blef_1)=rownames(blef_change_data)
blef_1_cor_mat=cor(scaled.blef_1)
# View(blef_1_cor_mat) #Uncomment to get table view in R
```

### Dissimilarity Matrix
```{r}
dist.mat = get_dist(scaled.blef_1) # Euclidean default

### Dissimilarity matrix
fviz_dist(dist.mat, gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07"))
```

### Optimal number of clusters
```{r}
### Optimal number of clusters?
fviz_nbclust(scaled.blef_1, kmeans, method = "wss") # 2 or 5?
fviz_nbclust(scaled.blef_1, kmeans, method = "silhouette") # 5

# Tibshirani, Walter, Hastie
# gap_stat = clusGap(scaled.blef_1, FUN = kmeans, K.max = 10, B = 100)
# fviz_gap_stat(gap_stat)
```

#### K-Means
```{r}
blef_1_kmeans=kmeans(scaled.blef_1, 5,nstart=25)
fviz_cluster(blef_1_kmeans, data = scaled.blef_1)
# Note: Why PF6 outlier? Check... (PF17 also seems outlierish)
```

#### Hierchical Clustering
```{r}
### Hierarchical clustering
blef_1_hclust1=hclust(dist.mat) #complete linkage
blef_1_hclust2=hclust(dist.mat,method='average')
blef_1_hclust3=hclust(dist.mat,method='single')
blef_1_hclust4 = hclust(dist.mat, method = "ward.D2" )
# Cut tree into 5 groups
grp = cutree(blef_1_hclust4, k = 5)

# Visualize
# par(mfrow=c(2,2),pty='m') # Uncomment to see on one plot to compare
plot(blef_1_hclust1, cex = 0.6) # plot tree complete linkage
rect.hclust(blef_1_hclust1, k = 5, border = 2:6) # add rectangle

plot(blef_1_hclust2, cex = 0.6) #average
rect.hclust(blef_1_hclust2, k = 5, border = 2:6) 

plot(blef_1_hclust3, cex = 0.6) #single
rect.hclust(blef_1_hclust3, k = 5, border = 2:6) 

plot(blef_1_hclust4, cex = 0.6) # Ward's method
rect.hclust(blef_1_hclust4, k = 5, border = 2:6)
```

#### Stability of clusters
```{r}
# Bootstrap to see stability of clusters
library(pvclust)
blef_1_pvclust1 = pvclust(scaled.blef_1)
blef_1_pvclust2 = pvclust(scaled.blef_1,
                              method.hclust="complete") 
# AU around 95% implies strongly supported by data
plot(blef_1_pvclust1)
pvrect(blef_1_pvclust1, alpha = 0.95)
plot(blef_1_pvclust2)
pvrect(blef_1_pvclust2, alpha = 0.95)
```

#### Most appropriate clustering method?
```{r}
# Choice of appropriate clustering algorithm (G Brock 2008)
library(clValid)
intern <- clValid(scaled.blef_1, nClust = 2:11,
                  clMethods = c("hierarchical","kmeans","pam"), 
                  validation = "internal")
summary(intern)

stab <- clValid(scaled.blef_1, nClust = 2:11,
                  clMethods = c("hierarchical","kmeans","pam"), 
                  validation = "stability")
optimalScores(stab)
summary(stab)
plot(stab)
```

### PCA
```{r}
### PCA to see how biomarkers are correlated
res.pca <- prcomp(scaled.blef_1)
# Visualize eigenvalues/variances
fviz_screeplot(res.pca, addlabels = TRUE, ylim = c(0, 50))
# Extract the results for variables
var = get_pca_var(res.pca)
# Contributions of variables to PC1
fviz_contrib(res.pca, choice = "var", axes = 1, top = 25)
# Contributions of variables to PC2
fviz_contrib(res.pca, choice = "var", axes = 2, top = 25)
# Control variable colors using their contributions to the principle axis
fviz_pca_var(res.pca, col.var="contrib",gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE # Avoid text overlapping
             ) + theme_minimal() + ggtitle("Variables - PCA")

### PCA without weight and BMI (only biomarkers)
res.pca <- prcomp(scaled.blef_1[,-1])
# Visualize eigenvalues/variances
fviz_screeplot(res.pca, addlabels = TRUE, ylim = c(0, 50))
# Extract the results for variables
var = get_pca_var(res.pca)
# Contributions of variables to PC1
fviz_contrib(res.pca, choice = "var", axes = 1, top = 25)
# Contributions of variables to PC2
fviz_contrib(res.pca, choice = "var", axes = 2, top = 25)
# Control variable colors using their contributions to the principle axis
fviz_pca_var(res.pca, col.var="contrib",gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE # Avoid text overlapping
             ) + theme_minimal() + ggtitle("Variables - PCA")
```
***

# EF $\to$ ER
```{r}
# EF to ER fractionalised-percentage changes 
# (Note: think about logs for skews ???)
efer_change=1+((fast_wide[Timepoint=="ER",-c(1,2)]-fast_wide[Timepoint=="EF",-c(1,2)])/fast_wide[Timepoint=="EF",-c(1,2)])
efer_change_data=select(efer_change,-matches("Weight|Waist|SBP|DBP|^Hb$|Ht|Neutr|Sodium|Potassium|Chloride|Calcium|ALT|Triglycerides|^HDL$"))
rownames(efer_change_data)=fast_wide[rownames(efer_change_data),"ID"]
# View(efer_change_data) #Uncomment to see table

# hist(efer_change_data[,"Weight"]) #Uncomment to see for skews/log-changes

(NAs_cols=which(colMeans(is.na(efer_change_data))!=0)) # 1 ER - Bilirubin only
```

## Version 1: Without Bilirubin (missing)
```{r}
## Version 1: Without Bilirubin
efer_1 = efer_change_data[,-NAs_cols]
```

### Scaling
```{r}
### TRIAL : CLUSTERING

scaled.efer_1= scale(efer_1)
rownames(scaled.efer_1)=rownames(efer_change_data)
efer_1_cor_mat=cor(scaled.efer_1)
# View(blef_1_cor_mat) #Uncomment to get table view in R
```

### Dissimilarity Matrix
```{r}
dist.mat = get_dist(scaled.efer_1) # Euclidean default

### Dissimilarity matrix
fviz_dist(dist.mat, gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07"))
```

### Optimal number of clusters
```{r}
### Optimal number of clusters?
fviz_nbclust(scaled.efer_1, kmeans, method = "wss") # 3 or 6?
fviz_nbclust(scaled.efer_1, kmeans, method = "silhouette") # 6

# Tibshirani, Walter, Hastie
# gap_stat = clusGap(scaled.blef_1, FUN = kmeans, K.max = 10, B = 100)
# fviz_gap_stat(gap_stat)
```

#### K-Means
```{r}
efer_1_kmeans=kmeans(scaled.efer_1, 6,nstart=25)
fviz_cluster(efer_1_kmeans, data = scaled.efer_1) # Good bit of overlap
# Note: Why PF6,PF10, PF14-15 outliers? Check...
```

#### Hierchical Clustering
```{r}
### Hierarchical clustering
efer_1_hclust1=hclust(dist.mat) #complete linkage
efer_1_hclust2=hclust(dist.mat,method='average')
efer_1_hclust3=hclust(dist.mat,method='single')
efer_1_hclust4 = hclust(dist.mat, method = "ward.D2" )
# Cut tree into 4 groups
grp = cutree(efer_1_hclust4, k = 6)

# Visualize
# par(mfrow=c(2,2),pty='m') # Uncomment to see on one plot to compare
plot(efer_1_hclust1, cex = 0.6) # plot tree complete linkage
rect.hclust(efer_1_hclust1, k = 6, border = 2:7) # add rectangle

plot(efer_1_hclust2, cex = 0.6) #average
rect.hclust(efer_1_hclust2, k = 6, border = 2:7) 

plot(efer_1_hclust3, cex = 0.6) #single
rect.hclust(efer_1_hclust3, k = 6, border = 2:7) 

plot(efer_1_hclust4, cex = 0.6) # Ward's method
rect.hclust(efer_1_hclust4, k = 6, border = 2:7)
```

#### Stability of clusters
```{r}
# Bootstrap to see stability of clusters
library(pvclust)
efer_1_pvclust1 = pvclust(scaled.efer_1)
efer_1_pvclust2 = pvclust(scaled.efer_1,
                              method.hclust="complete") 
# AU around 95% implies strongly supported by data
plot(efer_1_pvclust1)
pvrect(efer_1_pvclust1, alpha = 0.95)
plot(efer_1_pvclust2)
pvrect(efer_1_pvclust2, alpha = 0.95)
```

#### Most appropriate clustering method?
```{r}
# Choice of appropriate clustering algorithm (G Brock 2008)
library(clValid)
intern <- clValid(scaled.efer_1, nClust = 2:11,
                  clMethods = c("hierarchical","kmeans","pam"), 
                  validation = "internal")
summary(intern)
plot(intern)

stab <- clValid(scaled.efer_1, nClust = 2:11,
                  clMethods = c("hierarchical","kmeans","pam"), 
                  validation = "stability")
optimalScores(stab)
summary(stab)
plot(stab)
```

### PCA
```{r}
### PCA to see how biomarkers are correlated
res.pca <- prcomp(scaled.efer_1)
# Visualize eigenvalues/variances
fviz_screeplot(res.pca, addlabels = TRUE, ylim = c(0, 50))
# Extract the results for variables
var = get_pca_var(res.pca)
# Contributions of variables to PC1
fviz_contrib(res.pca, choice = "var", axes = 1, top = 25)
# Contributions of variables to PC2
fviz_contrib(res.pca, choice = "var", axes = 2, top = 25)
# Control variable colors using their contributions to the principle axis
fviz_pca_var(res.pca, col.var="contrib",gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE # Avoid text overlapping
             ) + theme_minimal() + ggtitle("Variables - PCA")

### PCA without weight and BMI (only biomarkers)
res.pca <- prcomp(scaled.efer_1[,-1])
# Visualize eigenvalues/variances
fviz_screeplot(res.pca, addlabels = TRUE, ylim = c(0, 50))
# Extract the results for variables
var = get_pca_var(res.pca)
# Contributions of variables to PC1
fviz_contrib(res.pca, choice = "var", axes = 1, top = 25)
# Contributions of variables to PC2
fviz_contrib(res.pca, choice = "var", axes = 2, top = 25)
# Control variable colors using their contributions to the principle axis
fviz_pca_var(res.pca, col.var="contrib",gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE # Avoid text overlapping
             ) + theme_minimal() + ggtitle("Variables - PCA")
```

```{r}
# Using ID as groups
# fast_wide_data=fast_wide[,-1]
# fast_wide_participants = fast_wide[,"ID"]

# segregate on timepoints

detach(fast_wide)
```
