---
title: "Subgroup correlations"
author: ''
date: "SJ"
output:
  html_document:
    toc: yes
    number_sections: true
    toc_depth: 5
  pdf_document:
    toc: yes
    number_sections: true
    toc_depth: 5
urlcolor: blue
linkcolor: red
---

```{r, global_options, include=FALSE}
rm(list = ls()) ##AM reset workspace - good practice to run clean
library(Cairo) # anti-aliasing for windows
library(GGally) # ggpairs plot
library(knitr) # used by kableExtra,rmarkdown
library(rmarkdown)
library(kableExtra)  #generate tables
library(pander)
library(arsenal)
require(survival)
library(naniar) # for visualising missing data
library(GGally) # for EDA
library(ggcorrplot) # for correlation matrix
library(plotly) # dynamic visualisations
library(ggplot2)
library(leaps) # Model selection
library(caret) # easy machine learning workflow
library(xgboost) # computing boosting algorithm
library(randomForest)
require(caTools) # split into train/test if required (for xgb etc.)
library(glmnet)
library(tidyverse) # clean and wrangle data; easy data manipulation and visualization

source("../SCRIPTS/Plot_functions_AM.R")

# Set working directory
# setwd("C:/Users/AM/Google Drive/_Healthy_Longevity_Group_G/Prolonged_Fasting_Study")
```


```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, warning=FALSE, message=FALSE)
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png")) #Windows only
theme_update(plot.title = element_text(hjust = 0.5)) # center plots
```

# Working Directory & Datasets
```{r, message=FALSE}

##AM please suppress trash outputs with chunk option: message=FALSE
##AM sometimes you need this to suppress stubborn outputs
#```{r results='hide', fig.keep='high'}

# fast_wide = read.csv("/DATA/Blood_changes_across_timepoints_rows_wide_AM.csv")
# fast_frac_wide = read.csv("/DATA/Blood_changes_across_timepoints_rows_fractions_AM.csv")

fast_wide = read.csv("../DATA/Blood_changes_across_timepoints_rows_wide_AM.csv",header=T,sep=",")
fast_frac_wide = read.csv("../DATA/Blood_changes_across_timepoints_rows_wide_fractions_AM.csv",header=T,sep=",")

# drop rows containing W1 timepoint - we will not use it
# fast_wide <- subset(fast_wide, Timepoint != "W1")
# fast_frac_wide <- subset(fast_frac_wide, Timepoint != "W1.f")

# working with fast_wide data
attach(fast_wide)
fast_wide$ID = as.factor(ID)
fast_wide$Timepoint=as.factor(Timepoint)
```

## Missing values?

```{r}
##AM add print() arguments to label each output
(total_missing=sum(is.na(fast_wide)))
(NAs_observations=which(rowMeans(is.na(fast_wide))!=0))
rowMeans(is.na(fast_wide[NAs_observations,])) # out of 50 observations
obs_totalnas=rowSums(is.na(fast_wide[NAs_observations,]))
names(obs_totalnas)=fast_wide[NAs_observations,"Timepoint"];obs_totalnas
# All NA-heavy ones W1 so no W1 analysis done.
# 3 BL (1st, 21st and 37th), 1 ER (40th obs) missing

plot(fast_wide$Timepoint,fast_wide$Triglycerides, xlab="Timepoint", ylab="Triglycerides")
```

# Datasets: BL, BL $\to$ EF, EF $\to$ ER

```{r}
# Baseline data
bl_data=fast_wide[Timepoint=="BL",]
rownames(bl_data)=bl_data$ID
bl_data=bl_data[,-c(1,2)]
baseline_data=dplyr::select(bl_data,-matches("Weight|Waist|SBP|DBP|^Hb$|Ht|Neutr|Sodium|Potassium|Chloride|Calcium|ALT|^HDL$"))
(NAs_cols=which(colMeans(is.na(baseline_data))!=0)) # Bililrubin and UrineSG
colSums(is.na(baseline_data[,NAs_cols]))
baseline_1=baseline_data[,-NAs_cols]
rm(bl_data)

# BL to EF fractionalised-percentage changes 
# (Note: think about logs for skews ???)
blef_change=1+((fast_wide[Timepoint=="EF",-c(1,2)]-fast_wide[Timepoint=="BL",-c(1,2)])/fast_wide[Timepoint=="BL",-c(1,2)])
blef_change_data=dplyr::select(blef_change,-matches("Weight|Waist|SBP|DBP|^Hb$|Ht|Neutr|Sodium|Potassium|Chloride|Calcium|ALT|^HDL$"))
rownames(blef_change_data)=fast_wide[rownames(blef_change_data),"ID"]
# View(blef_change_data) #Uncomment to see table
# hist(blef_change_data[,"Weight"]) #Uncomment to see for skews/log-changes
(NAs_cols=which(colMeans(is.na(blef_change_data))!=0)) # Bililrubin and UrineSG
colSums(is.na(blef_change_data[,NAs_cols]))
blef_1=blef_change_data[,-NAs_cols]
rm(blef_change_data,blef_change)

# EF to ER fractionalised-percentage changes 
# (Note: think about logs for skews ???)
efer_change=1+((fast_wide[Timepoint=="ER",-c(1,2)]-fast_wide[Timepoint=="EF",-c(1,2)])/fast_wide[Timepoint=="EF",-c(1,2)])
efer_change_data=dplyr::select(efer_change,-matches("Weight|Waist|SBP|DBP|^Hb$|Ht|Neutr|Sodium|Potassium|Chloride|Calcium|ALT|^HDL$"))
rownames(efer_change_data)=fast_wide[rownames(efer_change_data),"ID"]
# View(efer_change_data) #Uncomment to see table
# hist(efer_change_data[,"Weight"]) #Uncomment to see for skews/log-changes
(NAs_cols=which(colMeans(is.na(efer_change_data))!=0)) # 1 ER - Bilirubin only
efer_1 = efer_change_data[,-NAs_cols]
rm(efer_change_data,efer_change)

detach(fast_wide)
```

## Mean-Centering datasets
```{r}
# Mean-centering: https://link.springer.com/article/10.3758/s13428-015-0624-x
# Helps interecept interpretation, preserving slope and also helps a bit with multicollinearity
centr.blef_1=as.data.frame(scale(blef_1,center=T,scale=F))
centr.efer_1=as.data.frame(scale(efer_1,center=T,scale=F))
centr.baseline_1=as.data.frame(scale(baseline_1, center=T, scale=F))
```

## Important Functions

```{r}
# IMPORTANT FUNCTIONS
sig.corr_matrices = function(corr_mat, threshold = 0.7){
  corr_mat[abs(corr_mat)<threshold] = 0 # or 'NA' depending on visualisation used.
  return(corr_mat)
}


## Subgrouping function for delta fractional datasets
delta_frac_subgroups = function(delta_dataset, grp_var_name, same_dir=FALSE, thresh =0.5){
  attach(delta_dataset)
  if (same_dir) {
    # when biomarkers go in same direction for everyone
    high_grp=delta_dataset[abs(grp_var_name-1) >= thresh,]
    low_grp=delta_dataset[abs(grp_var_name-1) < thresh,] # here the group that doesn't go as high
    num_highr=dim(high_grp)[1]
    num_lowr=dim(low_grp)[1]
    output_list=list("higher_grp" = high_grp, "lower_grp" = low_grp,
                     "num_higher_grp" = num_highr,"num_lower_grp" = num_lowr)
    
    } else {
    incr_grp=delta_dataset[grp_var_name >= 1,]
    decr_grp=delta_dataset[grp_var_name < 1,]
    num_incr=dim(incr_grp)[1]
    num_decr=dim(decr_grp)[1]
    output_list=list("increased_grp" = incr_grp, "decreased_grp" = decr_grp,
                     "num_increased_grp" = num_incr,"num_decreased_grp" = num_decr)
    }
  detach(delta_dataset)
  return(output_list)
}

cor_scatter_plot= function(dataset,selected_cols=1:ncol(dataset), title=""){
  ggpairs(dataset, columns = selected_cols,
        lower = list(continuous ="smooth"), progress = F)+
    ggtitle(title)+
    theme(plot.title = element_text(hjust = 0.5))
}
```

### Opposite direction deltas

#### Glucose
```{r}
##Delta GLUCOSE##
# centr.blef_1

## Only BLEF
grps.blef=delta_frac_subgroups(blef_1,Glucose)
incr_grp=grps.blef$increased_grp; n_inc=grps.blef$num_increased_grp
decr_grp=grps.blef$decreased_grp; n_dec=grps.blef$num_decreased_grp

# Increased group
par(mfrow=c(2,1),pty="m")
corr_pearsons = sig.corr_matrices(cor(incr_grp,method="pearson"))
corr_pmat = cor_pmat(incr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BLEF, Grp: Gluc Inc") %>% ggplotly(width = 850, height = 750)

cor_scatter_plot(incr_grp, c("Testosterone", "IGFBP.1", "Beta.hydroxy.butyrate",
                            "IGF.1","BMI","IGFBP.2"), title="Data: BLEF, Grp: Gluc Inc")
# Checked and discarded FGF.21, Creatinine and AST
cor_scatter_plot(decr_grp, c("Testosterone", "IGFBP.1", "Beta.hydroxy.butyrate",
                            "IGF.1","BMI","IGFBP.2"), title="Data: BLEF, Grp: Gluc Dec")
cor_scatter_plot(centr.blef_1, c("Testosterone", "IGFBP.1", "Beta.hydroxy.butyrate",
                            "IGF.1","BMI","IGFBP.2"), title="Data: BLEF, Grp: All 20")



# Decreased group
corr_pearsons = sig.corr_matrices(cor(decr_grp,method="pearson"))
corr_pmat = cor_pmat(decr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BLEF, Grp: Gluc Dec") %>% ggplotly(width = 850, height = 750)
```


```{r, out.width="150%", out.height="150%"}
cor_scatter_plot(decr_grp, c("Leptin", "hsCRP", "Estradiol", "FGF.21", "SHBG",
                             "Protein","IGFBP.2","IGFBP.1","WBC","AST","Creatinine"), title="Data: BLEF, Grp: Gluc Dec")
cor_scatter_plot(incr_grp, c("Leptin", "hsCRP", "Estradiol", "FGF.21", "SHBG",
                             "Protein","IGFBP.2","IGFBP.1","WBC","AST","Creatinine"), title="Data: BLEF, Grp: Gluc Inc")
cor_scatter_plot(centr.blef_1, c("Leptin", "hsCRP", "Estradiol", "FGF.21", "SHBG",
                             "Protein","IGFBP.2","IGFBP.1","WBC","AST","Creatinine"), title="Data: BLEF, Grp: All 20")
```

```{r}
## With baseline
gluc_bl = cbind(blef_1$Glucose, baseline_1)
colnames(gluc_bl)[1]="Glucose.BLEF"

grps.bio_bl.blef=delta_frac_subgroups(gluc_bl, Glucose.BLEF)
incr_grp=grps.bio_bl.blef$increased_grp; n_inc=grps.bio_bl.blef$num_increased_grp
decr_grp=grps.bio_bl.blef$decreased_grp; n_dec=grps.bio_bl.blef$num_decreased_grp

# Increased group
par(mfrow=c(2,1),pty="m")
corr_pearsons = sig.corr_matrices(cor(incr_grp,method="pearson"))
corr_pmat = cor_pmat(incr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BL, Grp: Gluc_BLEF Inc") %>% ggplotly(width = 850, height = 750)
```


```{r, out.width="150%", out.height="150%"}
cor_scatter_plot(incr_grp, c("Glucose.BLEF", "IGFBP.2",
                             "SHBG", "Non_HDL_Chol","Beta.hydroxy.butyrate",
                             "UreaN2","IGFBP.1", "Leptin", "BMI", "FGF.21",
                             "Triglycerides","Insulin","AST"), 
                 title="Data: BL, Grp: Gluc_BLEF Inc")
#Leptin, BMI, FGF all significantly correlated with each other
cor_scatter_plot(decr_grp, c("Glucose.BLEF", "IGFBP.2",
                             "SHBG", "Non_HDL_Chol","Beta.hydroxy.butyrate",
                             "UreaN2","IGFBP.1", "Leptin", "BMI", "FGF.21",
                             "Triglycerides","Insulin","AST"), 
                 title="Data: BL, Grp: Gluc_BLEF Dec")
cor_scatter_plot(gluc_bl, c("Glucose.BLEF", "IGFBP.2",
                             "SHBG", "Non_HDL_Chol","Beta.hydroxy.butyrate",
                             "UreaN2","IGFBP.1", "Leptin", "BMI", "FGF.21",
                             "Triglycerides","Insulin","AST"), 
                 title="Data: BL, Grp: Gluc_BLEF All 20")
```

```{r}
# Decreased group
corr_pearsons = sig.corr_matrices(cor(decr_grp,method="pearson"))
corr_pmat = cor_pmat(decr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BL, Grp: Gluc_BLEF Dec") %>% ggplotly(width = 850, height = 750)
```


```{r, out.width="150%", out.height="150%"}
cor_scatter_plot(decr_grp, c("Protein","Glucose.BLEF", "hsCRP", 
                             "SHBG", "Triglycerides",
                             "LDL","IGFBP.2",'AST',
                             "Cholesterol", "Glucose"), 
                 title="Data: BL, Grp: Gluc_BLEF Dec")

#"LDL","IGFBP.2",'AST' correlated with each other
# "Protein","Glucose.BLEF", "hsCRP"
cor_scatter_plot(incr_grp, c("Protein","Glucose.BLEF", "hsCRP", 
                             "SHBG", "Triglycerides",
                             "LDL","IGFBP.2",'AST',
                             "Cholesterol", "Glucose"), 
                 title="Data: BL, Grp: Gluc_BLEF Inc")
cor_scatter_plot(gluc_bl, c("Protein","Glucose.BLEF", "hsCRP", 
                             "SHBG", "Triglycerides",
                             "LDL","IGFBP.2",'AST',
                             "Cholesterol", "Glucose"), 
                 title="Data: BL, Grp: Gluc_BLEF All 20")
```


```{r}

# Temporary break in code for testing
print("Early termination of code...")
knitr::knit_exit()

```



#### Insulin
```{r}
##Delta INSULIN##

## Only BLEF
grps.blef=delta_frac_subgroups(blef_1,Insulin)
incr_grp=grps.blef$increased_grp; n_inc=grps.blef$num_increased_grp
decr_grp=grps.blef$decreased_grp; n_dec=grps.blef$num_decreased_grp

# Increased group
par(mfrow=c(2,1),pty="m")
corr_pearsons = sig.corr_matrices(cor(incr_grp,method="pearson"))
corr_pmat = cor_pmat(incr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BLEF, Grp: Ins Inc") %>% ggplotly(width = 850, height = 750)

attach(incr_grp)
ggpairs(as.data.frame(cbind(Beta.hydroxy.butyrate, Leptin, 
                            BMI, IGF.1, IGFBP.1, Glucose, FreeT3)),
        lower = list(continuous ="smooth"), progress = F)+
  ggtitle("Data: BLEF, Grp: Ins Inc")+
  theme(plot.title = element_text(hjust = 0.5))
detach(incr_grp)

# Decreased group
corr_pearsons = sig.corr_matrices(cor(decr_grp,method="pearson"))
corr_pmat = cor_pmat(decr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BLEF, Grp: Ins Dec") %>% ggplotly(width = 850, height = 750)

attach(decr_grp)
ggpairs(as.data.frame(cbind(Insulin, Glucose,Beta.hydroxy.butyrate, FreeT3, Cortisol)),
        lower = list(continuous ="smooth"), progress = F)+
  ggtitle("Data: BLEF, Grp: Ins Dec")+
  theme(plot.title = element_text(hjust = 0.5))
detach(decr_grp)

## With baseline
ins_bl = cbind(blef_1$Insulin, baseline_1)
colnames(ins_bl)[1]="Insulin.BLEF"

grps.bio_bl.blef=delta_frac_subgroups(ins_bl, Insulin.BLEF)
incr_grp=grps.bio_bl.blef$increased_grp; n_inc=grps.bio_bl.blef$num_increased_grp
decr_grp=grps.bio_bl.blef$decreased_grp; n_dec=grps.bio_bl.blef$num_decreased_grp


# Increased group
par(mfrow=c(2,1),pty="m")
corr_pearsons = sig.corr_matrices(cor(incr_grp,method="pearson"))
corr_pmat = cor_pmat(incr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BL, Grp: Ins_BLEF Inc") %>% ggplotly(width = 850, height = 750)

attach(incr_grp)
ggpairs(as.data.frame(cbind(Insulin.BLEF, IGF.1,Cortisol,
                            Creatinine, hsCRP, Glucose, AST, Insulin,
                            LDL, Estradiol, WBC, SHBG)),
        lower = list(continuous ="smooth"), progress = F)+
  ggtitle("Data: BL, Grp: Ins_BLEF Inc")+
  theme(plot.title = element_text(hjust = 0.5))
detach(incr_grp)

# Decreased group
corr_pearsons = sig.corr_matrices(cor(decr_grp,method="pearson"))
corr_pmat = cor_pmat(decr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BL, Grp: Ins_BLEF Dec") %>% ggplotly(width = 850, height = 750)

attach(decr_grp)
ggpairs(as.data.frame(cbind(Insulin.BLEF, IGF.1, Cortisol, Testosterone, Estradiol)),
        lower = list(continuous ="smooth"), progress = F)+
  ggtitle("Data: BL, Grp: Ins_BLEF Dec")+
  theme(plot.title = element_text(hjust = 0.5))
detach(decr_grp)
```

#### Triglycerides
```{r}
##Delta TG##

## Only BLEF
grps.blef=delta_frac_subgroups(blef_1,Triglycerides)
incr_grp=grps.blef$increased_grp; n_inc=grps.blef$num_increased_grp
decr_grp=grps.blef$decreased_grp; n_dec=grps.blef$num_decreased_grp

# Increased group
par(mfrow=c(2,1),pty="m")
corr_pearsons = sig.corr_matrices(cor(incr_grp,method="pearson"))
corr_pmat = cor_pmat(incr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BLEF, Grp: Tg Inc") %>% ggplotly(width = 850, height = 750)

# Decreased group
corr_pearsons = sig.corr_matrices(cor(decr_grp,method="pearson"))
corr_pmat = cor_pmat(decr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BLEF, Grp: Tg Dec") %>% ggplotly(width = 850, height = 750)


## With baseline
tg_bl = cbind(blef_1$Triglycerides, baseline_1) # add Tg/HDL blef too if want correlations.
colnames(tg_bl)[1]="Triglycerides.BLEF"

grps.bio_bl.blef=delta_frac_subgroups(tg_bl, Triglycerides.BLEF)
incr_grp=grps.bio_bl.blef$increased_grp; n_inc=grps.bio_bl.blef$num_increased_grp
decr_grp=grps.bio_bl.blef$decreased_grp; n_dec=grps.bio_bl.blef$num_decreased_grp

# Increased group
par(mfrow=c(2,1),pty="m")
corr_pearsons = sig.corr_matrices(cor(incr_grp,method="pearson"))
corr_pmat = cor_pmat(incr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BL, Grp: Tg_BLEF Inc") %>% ggplotly(width = 850, height = 750)

# Decreased group
corr_pearsons = sig.corr_matrices(cor(decr_grp,method="pearson"))
corr_pmat = cor_pmat(decr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BL, Grp: Tg_BLEF Dec") %>% ggplotly(width = 850, height = 750)
```

#### LDL
```{r}
##Delta LDL##

## Only BLEF
grps.blef=delta_frac_subgroups(blef_1,LDL)
incr_grp=grps.blef$increased_grp; n_inc=grps.blef$num_increased_grp
decr_grp=grps.blef$decreased_grp; n_dec=grps.blef$num_decreased_grp

# Increased group
par(mfrow=c(2,1),pty="m")
corr_pearsons = sig.corr_matrices(cor(incr_grp,method="pearson"))
corr_pmat = cor_pmat(incr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BLEF, Grp: LDL Inc") %>% ggplotly(width = 850, height = 750)

# Decreased group
corr_pearsons = sig.corr_matrices(cor(decr_grp,method="pearson"))
corr_pmat = cor_pmat(decr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BLEF, Grp: LDL Dec") %>% ggplotly(width = 850, height = 750)


## With baseline
ldl_bl = cbind(blef_1$LDL, baseline_1)
colnames(ldl_bl)[1]="LDL.BLEF"

grps.bio_bl.blef=delta_frac_subgroups(ldl_bl, LDL.BLEF)
incr_grp=grps.bio_bl.blef$increased_grp; n_inc=grps.bio_bl.blef$num_increased_grp
decr_grp=grps.bio_bl.blef$decreased_grp; n_dec=grps.bio_bl.blef$num_decreased_grp

# Increased group
par(mfrow=c(2,1),pty="m")
corr_pearsons = sig.corr_matrices(cor(incr_grp,method="pearson"))
corr_pmat = cor_pmat(incr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BL, Grp: LDL_BLEF Inc") %>% ggplotly(width = 850, height = 750)

# Decreased group
corr_pearsons = sig.corr_matrices(cor(decr_grp,method="pearson"))
corr_pmat = cor_pmat(decr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BL, Grp: LDL_BLEF Dec") %>% ggplotly(width = 850, height = 750)
```

### Same direction deltas

#### IGF.1
```{r}
##Delta IGF.1##

## Only BLEF
grps.blef=delta_frac_subgroups(blef_1,IGF.1,same_dir = TRUE)
highr_grp=grps.blef$higher_grp; n_high=grps.blef$num_higher_grp
lowr_grp=grps.blef$lower_grp; n_low=grps.blef$num_lower_grp

# Higher group
par(mfrow=c(2,1),pty="m")
corr_pearsons = sig.corr_matrices(cor(highr_grp,method="pearson"))
corr_pmat = cor_pmat(highr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BLEF, Grp: IGF.1 High") %>% ggplotly(width = 850, height = 750)

# Lower group
corr_pearsons = sig.corr_matrices(cor(lowr_grp,method="pearson"))
corr_pmat = cor_pmat(lowr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BLEF, Grp: IGF.1 Low") %>% ggplotly(width = 850, height = 750)


## With baseline
igf1_bl = cbind(blef_1$IGF.1, baseline_1)
colnames(igf1_bl)[1]="IGF.1.BLEF"

grps.bio_bl.blef=delta_frac_subgroups(igf1_bl, IGF.1.BLEF, same_dir = TRUE)
highr_grp=grps.bio_bl.blef$higher_grp; n_high=grps.bio_bl.blef$num_higher_grp
lowr_grp=grps.bio_bl.blef$lower_grp; n_low=grps.bio_bl.blef$num_lower_grp

# Higher group
par(mfrow=c(2,1),pty="m")
corr_pearsons = sig.corr_matrices(cor(highr_grp,method="pearson"))
corr_pmat = cor_pmat(highr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BL, Grp: IGF.1_BLEF high") %>% ggplotly(width = 850, height = 750)

# lower group
corr_pearsons = sig.corr_matrices(cor(lowr_grp,method="pearson"))
corr_pmat = cor_pmat(lowr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BL, Grp: IGF.1_BLEF low") %>% ggplotly(width = 850, height = 750)
```

#### Cortisol
```{r}
##Delta Cortisol##

## Only BLEF
grps.blef=delta_frac_subgroups(blef_1,Cortisol,same_dir = TRUE,thresh=abs(median(Cortisol)-1))
highr_grp=grps.blef$higher_grp; n_high=grps.blef$num_higher_grp # maybe tweak thresh here
lowr_grp=grps.blef$lower_grp; n_low=grps.blef$num_lower_grp

# Higher group
par(mfrow=c(2,1),pty="m")
corr_pearsons = sig.corr_matrices(cor(highr_grp,method="pearson"))
corr_pmat = cor_pmat(highr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BLEF, Grp: Cort high") %>% ggplotly(width = 850, height = 750)

# lower group
corr_pearsons = sig.corr_matrices(cor(lowr_grp,method="pearson"))
corr_pmat = cor_pmat(lowr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BLEF, Grp: Cort low") %>% ggplotly(width = 850, height = 750)


## With baseline
cort_bl = cbind(blef_1$Cortisol, baseline_1)
colnames(cort_bl)[1]="Cortisol.BLEF"

grps.bio_bl.blef=delta_frac_subgroups(cort_bl, Cortisol.BLEF, same_dir = TRUE,thresh=(median(blef_1$Cortisol)-1))
highr_grp=grps.bio_bl.blef$higher_grp; n_high=grps.bio_bl.blef$num_higher_grp # maybe tweak thresh here
lowr_grp=grps.bio_bl.blef$lower_grp; n_low=grps.bio_bl.blef$num_lower_grp

# Higher group
par(mfrow=c(2,1),pty="m")
corr_pearsons = sig.corr_matrices(cor(highr_grp,method="pearson"))
corr_pmat = cor_pmat(highr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BL, Grp: Cort_BLEF high") %>% ggplotly(width = 850, height = 750)

# lower group
corr_pearsons = sig.corr_matrices(cor(lowr_grp,method="pearson"))
corr_pmat = cor_pmat(lowr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BL, Grp: Cort_BLEF low") %>% ggplotly(width = 850, height = 750)
```

#### FreeT3
```{r}
##Delta FreeT3##

## Only BLEF
grps.blef=delta_frac_subgroups(blef_1,FreeT3,same_dir = TRUE, thresh=abs(median(FreeT3)-1))
highr_grp=grps.blef$higher_grp; n_high=grps.blef$num_higher_grp # definitely tweak thresh here
lowr_grp=grps.blef$lower_grp; n_low=grps.blef$num_lower_grp

# Higher group
par(mfrow=c(2,1),pty="m")
corr_pearsons = sig.corr_matrices(cor(highr_grp,method="pearson"))
corr_pmat = cor_pmat(highr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BLEF, Grp: T3 high") %>% ggplotly(width = 850, height = 750)

# lower group
corr_pearsons = sig.corr_matrices(cor(lowr_grp,method="pearson"))
corr_pmat = cor_pmat(lowr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BLEF, Grp: T3 low") %>% ggplotly(width = 850, height = 750)


## With baseline
t3_bl = cbind(blef_1$FreeT3, baseline_1)
colnames(t3_bl)[1]="FreeT3.BLEF"

grps.bio_bl.blef=delta_frac_subgroups(t3_bl, FreeT3.BLEF, same_dir = TRUE,thresh=abs(median(FreeT3.BLEF)-1))
highr_grp=grps.bio_bl.blef$higher_grp; n_high=grps.bio_bl.blef$num_higher_grp
lowr_grp=grps.bio_bl.blef$lower_grp; n_low=grps.bio_bl.blef$num_lower_grp

# Higher group
par(mfrow=c(2,1),pty="m")
corr_pearsons = sig.corr_matrices(cor(highr_grp,method="pearson"))
corr_pmat = cor_pmat(highr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BL, Grp: T3_BLEF high") %>% ggplotly(width = 850, height = 750)

# lower group
corr_pearsons = sig.corr_matrices(cor(lowr_grp,method="pearson"))
corr_pmat = cor_pmat(lowr_grp, method='pearson')
ggcorrplot(corr_pearsons, p.mat = corr_pmat, type="lower",
           insig = "blank", sig.level = 0.05, title="Data: BL, Grp: T3_BLEF low") %>% ggplotly(width = 850, height = 750)
```