---
title: "PF Opp direction groups"
author: 'Significant biomarkers (good fits)'
date: "SJ"
output:
  html_document:
    toc: yes
    number_sections: true
    toc_depth: 5
  pdf_document:
    toc: yes
    number_sections: true
    toc_depth: 5
urlcolor: blue
linkcolor: red
---

```{r, global_options, include=FALSE}
rm(list = ls()) ##AM reset workspace - good practice to run clean
library(Cairo) # anti-aliasing for windows
library(GGally) # ggpairs plot
library(knitr) # used by kableExtra,rmarkdown
library(rmarkdown)
library(kableExtra)  #generate tables
library(pander)
library(arsenal)
require(survival)
library(naniar) # for visualising missing data
library(GGally) # for EDA
library(ggcorrplot) # for correlation matrix
library(plotly) # dynamic visualisations
library(ggplot2)
library(leaps) # Model selection
library(caret) # easy machine learning workflow
library(xgboost) # computing boosting algorithm
library(randomForest)
require(caTools) # split into train/test if required (for xgb etc.)
library(glmnet)
library(tidyverse) # clean and wrangle data; easy data manipulation and visualization
#library(ggpubr) #For gg single scatter plots (stat_cor)

# Set working directory
setwd("C:/Users/SJ/Desktop/Backup Mac/Desktop/STUDY/Usyd Statistician/Andrius Prolonged Fasting")
```


```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment=NA, warning=FALSE, message=FALSE)
theme_update(plot.title = element_text(hjust = 0.5)) # center plots
```

# Working Directory & Datasets
```{r, message=FALSE}

##AM please suppress trash outputs with chunk option: message=FALSE
##AM sometimes you need this to suppress stubborn outputs
#```{r results='hide', fig.keep='high'}

df_fast_wide = read.csv("Blood_changes_across_timepoints_rows_wide_AM.csv")
df_fast_frac_wide = read.csv("Blood_changes_across_timepoints_rows_fractions_AM.csv")

# df_fast_wide = read.csv("../DATA/Blood_changes_across_timepoints_rows_wide_AM.csv",header=T,sep=",")
# df_fast_frac_wide = read.csv("../DATA/Blood_changes_across_timepoints_rows_wide_fractions_AM.csv",header=T,sep=",")

# drop rows containing W1 timepoint - we will not use it
# df_fast_wide <- subset(df_fast_wide, Timepoint != "W1")
# df_fast_frac_wide <- subset(df_fast_frac_wide, Timepoint != "W1.f")

# working with df_fast_wide data
attach(df_fast_wide)
df_fast_wide$ID = as.factor(ID)
df_fast_wide$Timepoint=as.factor(Timepoint)
```

## Missing values?

```{r}
##AM add print() arguments to label each output
(total_missing=sum(is.na(df_fast_wide)))
(NAs_observations=which(rowMeans(is.na(df_fast_wide))!=0))
rowMeans(is.na(df_fast_wide[NAs_observations,])) # out of 50 observations
obs_totalnas=rowSums(is.na(df_fast_wide[NAs_observations,]))
names(obs_totalnas)=df_fast_wide[NAs_observations,"Timepoint"];obs_totalnas
# All NA-heavy ones W1 so no W1 analysis done.
# 3 BL (1st, 21st and 37th), 1 ER (40th obs) missing

plot(df_fast_wide$Timepoint,df_fast_wide$Triglycerides, xlab="Timepoint", ylab="Triglycerides")
```

# Datasets: BL, BL $\to$ EF, EF $\to$ ER

```{r, include=FALSE}
# Baseline data
bl_data=df_fast_wide[Timepoint=="BL",]
rownames(bl_data)=bl_data$ID
bl_data=bl_data[,-c(1,2)]
df_baseline_data=dplyr::select(bl_data,-matches("Weight|Waist|SBP|DBP|^Hb$|Ht|Neutr|Sodium|Potassium|Chloride|Calcium|ALT|^HDL$"))
(NAs_cols=which(colMeans(is.na(df_baseline_data))!=0)) # Bililrubin and UrineSG
colSums(is.na(df_baseline_data[,NAs_cols]))
df_baseline_1=df_baseline_data[,-NAs_cols]
rm(bl_data)

# BL to EF fractionalised-percentage changes 
# (Note: think about logs for skews ???)
blef_change=1+((df_fast_wide[Timepoint=="EF",-c(1,2)]-df_fast_wide[Timepoint=="BL",-c(1,2)])/df_fast_wide[Timepoint=="BL",-c(1,2)])
blef_change_data=dplyr::select(blef_change,-matches("Weight|Waist|SBP|DBP|^Hb$|Ht|Neutr|Sodium|Potassium|Chloride|Calcium|ALT|^HDL$"))
rownames(blef_change_data)=df_fast_wide[rownames(blef_change_data),"ID"]
# View(blef_change_data) #Uncomment to see table
# hist(blef_change_data[,"Weight"]) #Uncomment to see for skews/log-changes
(NAs_cols=which(colMeans(is.na(blef_change_data))!=0)) # Bililrubin and UrineSG
colSums(is.na(blef_change_data[,NAs_cols]))
df_blef_1=blef_change_data[,-NAs_cols]
rm(blef_change_data,blef_change)

# EF to ER fractionalised-percentage changes 
# (Note: think about logs for skews ???)
efer_change=1+((df_fast_wide[Timepoint=="ER",-c(1,2)]-df_fast_wide[Timepoint=="EF",-c(1,2)])/df_fast_wide[Timepoint=="EF",-c(1,2)])
efer_change_data=dplyr::select(efer_change,-matches("Weight|Waist|SBP|DBP|^Hb$|Ht|Neutr|Sodium|Potassium|Chloride|Calcium|ALT|^HDL$"))
rownames(efer_change_data)=df_fast_wide[rownames(efer_change_data),"ID"]
# View(efer_change_data) #Uncomment to see table
# hist(efer_change_data[,"Weight"]) #Uncomment to see for skews/log-changes
(NAs_cols=which(colMeans(is.na(efer_change_data))!=0)) # 1 ER - Bilirubin only
df_efer_1 = efer_change_data[,-NAs_cols]
rm(efer_change_data,efer_change)

detach(df_fast_wide)
```

## Mean-Centering datasets
```{r}
# Mean-centering: https://link.springer.com/article/10.3758/s13428-015-0624-x
# Helps interecept interpretation, preserving slope and also helps a bit with multicollinearity
centr.df_blef_1=as.data.frame(scale(df_blef_1,center=T,scale=F))
centr.df_efer_1=as.data.frame(scale(df_efer_1,center=T,scale=F))
centr.df_baseline_1=as.data.frame(scale(df_baseline_1, center=T, scale=F))
```

## Important Functions

```{r, include=FALSE}
# IMPORTANT FUNCTIONS

## Assigns 0 to correlations under threshold for easy visualisation of strong correlations
func_sig.corr_matrices = function(corr_mat, threshold = 0.7){
  corr_mat[abs(corr_mat)<threshold] = 0 # or 'NA' depending on visualisation used.
  return(corr_mat)
}

## Subgrouping function for delta fractional and delta basic datasets
func_delta_frac_subgroups = function(delta_dataset, grp_var_name,
                                same_dir=FALSE, basic_diff=FALSE,thresh =0.5) {
  attach(delta_dataset)
  if (basic_diff){
    incr_grp=delta_dataset[grp_var_name >= 0,]
    decr_grp=delta_dataset[grp_var_name < 0,]
    num_incr=dim(incr_grp)[1]
    num_decr=dim(decr_grp)[1]
    output_list=list("increased_grp" = incr_grp, "decreased_grp" = decr_grp,
                     "num_increased_grp" = num_incr,"num_decreased_grp" = num_decr)
  } else {
    if (same_dir) {
    # when biomarkers go in same direction for everyone
    high_grp=delta_dataset[abs(grp_var_name-1) >= thresh,]
    low_grp=delta_dataset[abs(grp_var_name-1) < thresh,] # here the group that doesn't go as high
    num_highr=dim(high_grp)[1]
    num_lowr=dim(low_grp)[1]
    output_list=list("higher_grp" = high_grp, "lower_grp" = low_grp,
                     "num_higher_grp" = num_highr,"num_lower_grp" = num_lowr)
    
    } else {
    incr_grp=delta_dataset[grp_var_name >= 1,]
    decr_grp=delta_dataset[grp_var_name < 1,]
    num_incr=dim(incr_grp)[1]
    num_decr=dim(decr_grp)[1]
    output_list=list("increased_grp" = incr_grp, "decreased_grp" = decr_grp,
                     "num_increased_grp" = num_incr,"num_decreased_grp" = num_decr)
    }
  }
  detach(delta_dataset)
  return(output_list)
}

## ggScatter plot of selected biomarkers
func_cor_scatter_plot= function(dataset,selected_cols=1:ncol(dataset), title=""){
  ggpairs(dataset, columns = selected_cols,
          lower = list(continuous ="smooth"), progress = F)+
    ggtitle(title)+
    theme(plot.title = element_text(hjust = 0.5))
}

# Single scatter plot
# sp <- ggscatter(df, x = "wt", y = "mpg",
#    add = "reg.line",  # Add regressin line
#    add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
#    conf.int = TRUE # Add confidence interval
#    )
# # Add correlation coefficient
# sp + stat_cor(method = "pearson", p.accuracy=0.001, r.accuracy=0.001)

## More strict than sig.corr_matrics by only looking at significant p-values.
func_cor_pval_plot = function(dataset,method="spearman", thresh = 0.7, sig_level=0.05,title=""){
  corr_matrix = func_sig.corr_matrices(cor(dataset,method=method), threshold = thresh)
  corr_pmat = cor_pmat(dataset, method=method)
  ggcorrplot(corr_matrix, p.mat = corr_pmat, type="lower",
             insig = "blank", sig.level = sig_level, title=title) %>% ggplotly(width=850, height=750)
}


## For descriptive statistics (see below)
func_Mode = function(x) {
  ux = unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

func_neat_descriptive_stats = function(datatable,title="", sig.figs=3){
  
  stats_table=data.frame(round(cbind(apply(datatable,2,mean), apply(datatable,2,sd),
                                     apply(datatable,2,Mode),
                                     t(apply(datatable,2,quantile))), sig.figs))
  stats_table=pander(stats_table, keep.line.breaks = TRUE,style="grid",caption=title,
                     col.names=c("Mean", "Std. Dev", "Mode",
                                 "Min", "1st Q", "Median\n (2nd Q)",
                                 "3rd Q","Max"),
                     row.names=c("Body weight", "LDL Cholestrol\n (mg/dL)", 
                                 "HDL Cholestrol\n (mg/dL)",
                                 "Triglyceride, Tg\n (mg/dL)",
                                 "Glucose-Fasting\n (mg/dL)", 
                                 "Insulin-Fasting\n (uIU/mL)",
                                 "C-Reactive Protein\n (ug/mL)",
                                 "TNF-alpha \n (pg/mL)",
                                 "Triiodothyronine, T3\n (ng/dL)", 
                                 "Cortisol\n (ug/dL)", "IGF-1\n (ng/mL)",
                                 "IGFBP-1\n (pg/mL)", "IGFBP-3\n (ng/mL)"),
                     split.table=Inf)
  return (stats_table)
}

func_elbow_test = function(criterion){
  attach(criterion)
  plot(1:dim(criterion[1]),AdjR2*100,pch=20,type='b',main="Elbow test", xlab="Number of explanatory variables", ylab="AdjR2")
  plot(1:dim(exhaustive_models[1]),Cp, pch=20, type="b", col="red",
       main="Elbow test", xlab="Number of explanatory variables", ylab="Mallow's Cp")
  plot(1:dim(exhaustive_models[1]),BIC,  pch=20, type="b", col="blue",
       main="Elbow test", xlab="Number of explanatory variables", ylab="BIC")
  detach(criterion)
  
}

func_resids_diagnostics = function(fit,explanatory_vars,title="", ident=FALSE,
                              subtitle="", cooks_title=""){
  par(mfcol=c(2,2),oma = c(2,0,1,0))
  n = dim(explanatory_vars)[1]

  plot(c(1:n),hat(explanatory_vars),ylim=c(0,max(hat(w))), 
       main="(a) Leverage for datapoints", 
       xlab="Observation Number",
       ylab="Leverage",pch=20)
  # segments(c(1:n),0,c(1:n),hat(w))
  abline(h=(2*3)/n,col="red") #2(p+1)/n
  
  # par(pty="s")
  p2=qqnorm(residuals(fit),
            main="(b) Quantile-Quantile Plot",
            xlab="Gaussian Quantiles",
            ylab="Residuals",pch=20)
  qqline(residuals(fit), col="red")
  abline(v=0,col="Blue",lty=2)
  abline(h=0,col="Blue",lty=2)
  if (ident){
    identify(p2, plot = TRUE)
  }
  
  # par(pty="m")
  
  plot(fitted(fit),residuals(fit),
       main="(c) Residual Plot",
       xlab="Fitted Values",
       ylab="Residuals",pch=20)
  abline(h=0,col="blue",lty=2)
  abline(h=c(1.96,-1.96),lty=2,col="red")
  lines(lowess(fitted(fit),residuals(fit)), col="red")
  # identify(fitted(fit),residuals(fit))
  
  plot(fitted(fit),abs(residuals(fit)),
       main="(d) Absolute Residual Plot",
       xlab="Fitted Values",
       ylab="Absolute Residuals",pch=20)
  lines(lowess(fitted(fit),abs(residuals(fit))), col="red")
  # identify(fitted(fit),abs(residuals(fit)))
  mtext(title, col = "red",
        side=3,line=-1.5,outer=T,cex=1.2,font=2)
  mtext(subtitle,
        side=1,line=1,outer=T,cex=1.1,col='blue')
  
  par(mfrow=c(1,1), pty="m")
  plot(cooks.distance(fit),pch=20, 
       main=cooks_title, 
       col.main="red", ylab="Cook's Distance", xlab="Observation")
  abline(h=0.01,col='blue', lty=2)
  mtext(subtitle,
        side=1,line=0.5,outer=T,cex=1.1,col='blue')
  if (ident){
    identify(cooks.distance(fit), plot = TRUE, col='red')
  }
}
```

### Opposite direction deltas

#### Glucose
```{r}
##Delta GLUCOSE##

## Only BLEF
grps.blef=func_delta_frac_subgroups(df_blef_1,Glucose)
incr_grp=grps.blef$increased_grp; n_inc=grps.blef$num_increased_grp
decr_grp=grps.blef$decreased_grp; n_dec=grps.blef$num_decreased_grp

# Increased group
func_cor_scatter_plot(incr_grp, c("BMI","Glucose","IGF.1"), title="Data: BLEF, Grp: Gluc Inc (Part I)")

func_cor_scatter_plot(incr_grp, c("Beta.hydroxy.butyrate","IGFBP.1","IGF.1"), title="Data: BLEF, Grp: Gluc Inc (Part II)")

func_cor_scatter_plot(incr_grp, c("IGFBP.1", "Testosterone","SHBG"), title="Data: BLEF, Grp: Gluc Inc (Part III)")

func_cor_scatter_plot(incr_grp, c("IGFBP.2", "SHBG"), title="Data: BLEF, Grp: Gluc Inc (Part IV)")

# Decreased group
func_cor_scatter_plot(decr_grp, c("Leptin", "hsCRP"), title="Data: BLEF, Grp: Gluc Dec (I)")
func_cor_scatter_plot(decr_grp, c("Leptin", "Estradiol"), title="Data: BLEF, Grp: Gluc Dec (II)")
func_cor_scatter_plot(decr_grp, c("IGFBP.1", "hsCRP"), title="Data: BLEF, Grp: Gluc Dec (III)")

## With baseline
gluc_bl = cbind(df_blef_1$Glucose, df_baseline_1)
colnames(gluc_bl)[1]="Glucose.BLEF"

grps.bio_bl.blef=func_delta_frac_subgroups(gluc_bl, Glucose.BLEF)
incr_grp=grps.bio_bl.blef$increased_grp; n_inc=grps.bio_bl.blef$num_increased_grp
decr_grp=grps.bio_bl.blef$decreased_grp; n_dec=grps.bio_bl.blef$num_decreased_grp

# Increased group

func_cor_scatter_plot(incr_grp, c("FGF.21","Leptin","BMI"), 
                 title="Data: BL, Grp: Gluc_BLEF Inc (Part I)")

func_cor_scatter_plot(incr_grp, c("FGF.21","AST","Triglycerides"), 
                 title="Data: BL, Grp: Gluc_BLEF Inc (Part II)")

func_cor_scatter_plot(incr_grp, c("Insulin","Triglycerides"), 
                 title="Data: BL, Grp: Gluc_BLEF Inc (Part III)")


# Decreased group
func_cor_scatter_plot(decr_grp, c("Glucose.BLEF", "hsCRP", "Protein"), 
                 title="Data: BL, Grp: Gluc_BLEF Dec (Part I)")

func_cor_scatter_plot(decr_grp, c("Triglycerides","SHBG", "Trig_HDL_Ratio"), 
                 title="Data: BL, Grp: Gluc_BLEF Dec (Part II)")

func_cor_scatter_plot(decr_grp, c("IGFBP.2","Adiponectin","SHBG"), 
                 title="Data: BL, Grp: Gluc_BLEF Dec (Part III)")

func_cor_scatter_plot(decr_grp, c("IGFBP.2","AST"), 
                 title="Data: BL, Grp: Gluc_BLEF Dec (Part IV)")
```

#### Insulin
```{r}
##Delta INSULIN##

## Only BLEF
grps.blef=func_delta_frac_subgroups(df_blef_1,Insulin)
incr_grp=grps.blef$increased_grp; n_inc=grps.blef$num_increased_grp
decr_grp=grps.blef$decreased_grp; n_dec=grps.blef$num_decreased_grp

# Increased group
func_cor_scatter_plot(incr_grp, c("Leptin", "BMI", "IGF.1"), 
                 title="Data: BLEF, Grp: Ins Inc (Part I)")

func_cor_scatter_plot(incr_grp, c("Beta.hydroxy.butyrate","IGFBP.1"), 
                 title="Data: BLEF, Grp: Ins Inc (Part II)")

func_cor_scatter_plot(incr_grp, c("BMI","IGFBP.2", "IGF.1"), 
                 title="Data: BLEF, Grp: Ins Inc (Part III)")

# Decreased group

func_cor_scatter_plot(decr_grp, c("Glucose","Beta.hydroxy.butyrate", "FreeT3"), 
                 title="Data: BLEF, Grp: Ins Dec (Part I)")

func_cor_scatter_plot(decr_grp, c("Insulin","FreeT3"), 
                 title="Data: BLEF, Grp: Ins Dec (Part II)")

func_cor_scatter_plot(decr_grp, c("AlkPhos.","Beta.hydroxy.butyrate"), 
                 title="Data: BLEF, Grp: Ins Dec (Part III)")

## With baseline
ins_bl = cbind(df_blef_1$Insulin, df_baseline_1)
colnames(ins_bl)[1]="Insulin.BLEF"

grps.bio_bl.blef=func_delta_frac_subgroups(ins_bl, Insulin.BLEF)
incr_grp=grps.bio_bl.blef$increased_grp; n_inc=grps.bio_bl.blef$num_increased_grp
decr_grp=grps.bio_bl.blef$decreased_grp; n_dec=grps.bio_bl.blef$num_decreased_grp

# Increased group
func_cor_scatter_plot(incr_grp, c("Glucose","AST"), 
                 title="Data: BL, Grp: Ins_BLEF Inc (Part I)")

func_cor_scatter_plot(incr_grp, c("Insulin.BLEF","IGF.1"), 
                 title="Data: BL, Grp: Ins_BLEF Inc (Part II)")

func_cor_scatter_plot(incr_grp, c("Glucose","Insulin"),
                 title="Data: BL, Grp: Ins_BLEF Inc (Part III)")

# Decreased group
func_cor_scatter_plot(decr_grp, c("FGF.21", "BMI", "Leptin"), 
                 title="Data: BL, Grp: Ins_BLEF Dec (Part I)")

func_cor_scatter_plot(decr_grp, c("Triglycerides", "Insulin"), 
                 title="Data: BL, Grp: Ins_BLEF Dec (Part II)")
```

#### Triglycerides
```{r}
##Delta TG##

## Only BLEF
grps.blef=func_delta_frac_subgroups(df_blef_1,Triglycerides)
incr_grp=grps.blef$increased_grp; n_inc=grps.blef$num_increased_grp
decr_grp=grps.blef$decreased_grp; n_dec=grps.blef$num_decreased_grp

# Increased group
func_cor_scatter_plot(incr_grp, c("Beta.hydroxy.butyrate", "Glucose",
                             "Insulin"),
                 title="Data: BLEF, Grp: Tg Inc (Part I)")

func_cor_scatter_plot(incr_grp, c("Glucose", "WBC"),
                 title="Data: BLEF, Grp: Tg Inc (Part II)")

func_cor_scatter_plot(incr_grp, c("Triglycerides", "AST", "Trig_HDL_Ratio"),
                 title="Data: BLEF, Grp: Tg Inc (III)")


# Decreased group
func_cor_scatter_plot(decr_grp, c("Beta.hydroxy.butyrate", "Cholesterol", "LDL"),
                 title="Data: BLEF, Grp: Tg Dec (Part I)")

func_cor_scatter_plot(decr_grp, c("Trig_HDL_Ratio","IGFBP.3","IGFBP.2"),
                 title="Data: BLEF, Grp: Tg Dec (Part II)")

func_cor_scatter_plot(decr_grp, c("LDL", "Glucose"),
                 title="Data: BLEF, Grp: Tg Dec (Part III)")

func_cor_scatter_plot(decr_grp, c("Triglycerides", "Cortisol"),
                 title="Data: BLEF, Grp: Tg Dec (Part IV)")

func_cor_scatter_plot(decr_grp, c("Cortisol", "hsCRP"),
                 title="Data: BLEF, Grp: Tg Dec (Part V)")

func_cor_scatter_plot(decr_grp, c("Triglycerides", "Creatinine"),
                 title="Data: BLEF, Grp: Tg Dec (Part VI)")

func_cor_scatter_plot(decr_grp, c("IGFBP.1", "Creatinine"),
                 title="Data: BLEF, Grp: Tg Dec (Part VII)")

func_cor_scatter_plot(decr_grp, c("Trig_HDL_Ratio","AST"),
                 title="Data: BLEF, Grp: Tg Dec (Part VIII)")

func_cor_scatter_plot(decr_grp, c("WBC", "UreaN2"),
                 title="Data: BLEF, Grp: Tg Dec (Part IX)")

func_cor_scatter_plot(decr_grp, c("Insulin", "UreaN2"),
                 title="Data: BLEF, Grp: Tg Dec (Part X)")

func_cor_scatter_plot(decr_grp, c("LDL", "Insulin"),
                 title="Data: BLEF, Grp: Tg Dec (Part XI)")

## With baseline
tg_bl = cbind(df_blef_1$Triglycerides, df_baseline_1)
colnames(tg_bl)[1]="Tg.BLEF"

grps.bio_bl.blef=func_delta_frac_subgroups(tg_bl, Tg.BLEF)
incr_grp=grps.bio_bl.blef$increased_grp; n_inc=grps.bio_bl.blef$num_increased_grp
decr_grp=grps.bio_bl.blef$decreased_grp; n_dec=grps.bio_bl.blef$num_decreased_grp


# Increased group
func_cor_scatter_plot(incr_grp, c("Leptin","BMI","FGF.21"),
                 title="Data: BL, Grp: Tg_BLEF Inc (Part I)")

func_cor_scatter_plot(incr_grp, c("Glucose", "Beta.hydroxy.butyrate","hsCRP"),
                 title="Data: BL, Grp: Tg_BLEF Inc (Part II)")

func_cor_scatter_plot(incr_grp, c("UricAcid","hsCRP"),
                 title="Data: BL, Grp: Tg_BLEF Inc (Part III)")

# Decreased group
func_cor_scatter_plot(decr_grp, c("hsCRP", "Mono"),
                 title="Data: BL, Grp: Tg_BLEF Dec (Part I)")

func_cor_scatter_plot(decr_grp, c("FGF.21", "IGFBP.1", "Tg.BLEF"),
                 title="Data: BL, Grp: Tg_BLEF Dec (Part II)")

func_cor_scatter_plot(decr_grp, c("Leptin", "BMI", "FGF.21"),
                 title="Data: BL, Grp: Tg_BLEF Dec (Part III)")

func_cor_scatter_plot(decr_grp, c("Beta.hydroxy.butyrate", "CO2", "BMI"),
                 title="Data: BL, Grp: Tg_BLEF Dec (Part IV)")

func_cor_scatter_plot(decr_grp, c("Lymp", "CO2"),
                 title="Data: BL, Grp: Tg_BLEF Dec (Part V)")

func_cor_scatter_plot(decr_grp, c("LDL", "Triglycerides", "UricAcid"),
                 title="Data: BL, Grp: Tg_BLEF Dec (Part VI)")

func_cor_scatter_plot(decr_grp, c("Adiponectin", "IGFBP.1", "Cortisol"),
                 title="Data: BL, Grp: Tg_BLEF Dec (Part VII)")

func_cor_scatter_plot(decr_grp, c("Cholesterol", "IGFBP.2"),
                 title="Data: BL, Grp: Tg_BLEF Dec (Part VIII)")

func_cor_scatter_plot(decr_grp, c("WBC", "Cortisol"),
                 title="Data: BL, Grp: Tg_BLEF Dec (Part IX)")

func_cor_scatter_plot(decr_grp, c("Cholesterol", "IGFBP.2"),
                 title="Data: BL, Grp: Tg_BLEF Dec (Part X)")
```

#### LDL
```{r}
##Delta LDL##

## Only BLEF
grps.blef=func_delta_frac_subgroups(df_blef_1,LDL)
incr_grp=grps.blef$increased_grp; n_inc=grps.blef$num_increased_grp
decr_grp=grps.blef$decreased_grp; n_dec=grps.blef$num_decreased_grp

# Increased group
func_cor_scatter_plot(incr_grp, c("Beta.hydroxy.butyrate", "Glucose", "Insulin"),
                 title="Data: BLEF, Grp: LDL Inc (Part I)")

func_cor_scatter_plot(incr_grp, c("FGF.21","Mono","AlkPhos."),
                 title="Data: BLEF, Grp: LDL Inc (Part II)")

func_cor_scatter_plot(incr_grp, c("Cholesterol","Non_HDL_Chol","IGFBP.3"),
                 title="Data: BLEF, Grp: LDL Inc (Part III)")

func_cor_scatter_plot(incr_grp, c("Beta.hydroxy.butyrate", "Triglycerides"),
                 title="Data: BLEF, Grp: LDL Inc (Part IV)")

func_cor_scatter_plot(incr_grp, c("BMI", "Mono"),
                 title="Data: BLEF, Grp: LDL Inc (Part V)")


# Decreased group
func_cor_scatter_plot(decr_grp, c("IGFBP.1","Beta.hydroxy.butyrate", "Albumin"),
                 title="Data: BLEF, Grp: LDL Dec (Part I)")

func_cor_scatter_plot(decr_grp, c("UreaN2", "IGFBP.3"),
                 title="Data: BLEF, Grp: LDL Dec (Part II)")

func_cor_scatter_plot(decr_grp, c("LDL", "Triglycerides"),
                 title="Data: BLEF, Grp: LDL Dec (Part III)")

## With baseline
ldl_bl = cbind(df_blef_1$LDL, df_baseline_1)
colnames(ldl_bl)[1]="LDL.BLEF"

grps.bio_bl.blef=func_delta_frac_subgroups(ldl_bl, LDL.BLEF)
incr_grp=grps.bio_bl.blef$increased_grp; n_inc=grps.bio_bl.blef$num_increased_grp
decr_grp=grps.bio_bl.blef$decreased_grp; n_dec=grps.bio_bl.blef$num_decreased_grp


# Increased group

func_cor_scatter_plot(incr_grp, c("Insulin", "Leptin","BMI","FGF.21"),
                 title="Data: BL, Grp: LDL_BLEF Inc (Part I)")

func_cor_scatter_plot(incr_grp, c("IGFBP.1","IGFBP.2","Beta.hydroxy.butyrate"),
                 title="Data: BL, Grp: LDL_BLEF Inc (Part II)")

func_cor_scatter_plot(incr_grp, c("Creatinine", "LDL"),
                 title="Data: BL, Grp: LDL_BLEF Inc (Part III)")

func_cor_scatter_plot(incr_grp, c("IGFBP.2","CO2"),
                 title="Data: BL, Grp: LDL_BLEF Inc (Part IV)")

func_cor_scatter_plot(incr_grp, c("Triglycerides", "SHBG","Trig_HDL_Ratio"),
                 title="Data: BL, Grp: LDL_BLEF Inc (Part V)")


# Decreased group
func_cor_scatter_plot(decr_grp, c("Insulin","FGF.21"),
                 title="Data: BL, Grp: LDL_BLEF Dec (Part I)")

func_cor_scatter_plot(decr_grp, c("LDL", "SHBG"),
                 title="Data: BL, Grp: LDL_BLEF Dec (Part II)")

func_cor_scatter_plot(decr_grp, c("BMI","FGF.21","AST"),
                 title="Data: BL, Grp: LDL_BLEF Dec (Part III)")

func_cor_scatter_plot(decr_grp, c("Mono","Protein","IGFBP.3"),
                 title="Data: BL, Grp: LDL_BLEF Dec (Part IV)")

func_cor_scatter_plot(decr_grp, c("Cortisol", "Insulin","Triglycerides"),
                 title="Data: BL, Grp: LDL_BLEF Dec (Part V)")

func_cor_scatter_plot(decr_grp, c("Adiponectin", "Albumin"),
                 title="Data: BL, Grp: LDL_BLEF Dec (Part VI)")
```