---
title: "PF Opp direction groups"
author: 'Candidate biomarkers (Fig.3b)'
date: "SJ"
output:
  pdf_document:
    toc: yes
    number_sections: true
    toc_depth: 5
  html_document:
    toc: yes
    number_sections: true
    toc_depth: 5
urlcolor: blue
linkcolor: red
---

```{r, global_options, include=FALSE}
rm(list = ls()) ##AM reset workspace - good practice to run clean
library(Cairo) # anti-aliasing for windows
library(GGally) # ggpairs plot
library(knitr) # used by kableExtra,rmarkdown
library(rmarkdown)
library(kableExtra)  #generate tables
library(pander)
library(arsenal)
require(survival)
library(naniar) # for visualising missing data
library(GGally) # for EDA
library(ggcorrplot) # for correlation matrix
library(plotly) # dynamic visualisations
library(ggplot2)
library(leaps) # Model selection
library(caret) # easy machine learning workflow
library(xgboost) # computing boosting algorithm
library(randomForest)
require(caTools) # split into train/test if required (for xgb etc.)
library(glmnet)
library(tidyverse) # clean and wrangle data; easy data manipulation and visualization
#library(ggpubr) #For gg single scatter plots (stat_cor)

source("C:/Users/SJ/Desktop/Backup Mac/Desktop/STUDY/Usyd Statistician/Functions/Subgroup_corr_resids.R") # required functions

# Set working directory
setwd("C:/Users/SJ/Desktop/Backup Mac/Desktop/STUDY/Usyd Statistician/Andrius Prolonged Fasting")
```


```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment=NA, warning=FALSE, message=FALSE)
theme_update(plot.title = element_text(hjust = 0.5)) # center plots
```

# Working Directory & Datasets
```{r, message=FALSE}

##AM please suppress trash outputs with chunk option: message=FALSE
##AM sometimes you need this to suppress stubborn outputs
#```{r results='hide', fig.keep='high'}

df_fast_wide = read.csv("Blood_changes_across_timepoints_rows_wide_AM.csv")
df_fast_frac_wide = read.csv("Blood_changes_across_timepoints_rows_fractions_AM.csv")

# df_fast_wide = read.csv("../DATA/Blood_changes_across_timepoints_rows_wide_AM.csv",header=T,sep=",")
# df_fast_frac_wide = read.csv("../DATA/Blood_changes_across_timepoints_rows_wide_fractions_AM.csv",header=T,sep=",")

# drop rows containing W1 timepoint - we will not use it
# df_fast_wide <- subset(df_fast_wide, Timepoint != "W1")
# df_fast_frac_wide <- subset(df_fast_frac_wide, Timepoint != "W1.f")

# working with df_fast_wide data
attach(df_fast_wide)
df_fast_wide$ID = as.factor(ID)
df_fast_wide$Timepoint=as.factor(Timepoint)
```

## Missing values?

```{r}
##AM add print() arguments to label each output
(total_missing=sum(is.na(df_fast_wide)))
(NAs_observations=which(rowMeans(is.na(df_fast_wide))!=0))
rowMeans(is.na(df_fast_wide[NAs_observations,])) # out of 50 observations
obs_totalnas=rowSums(is.na(df_fast_wide[NAs_observations,]))
names(obs_totalnas)=df_fast_wide[NAs_observations,"Timepoint"];obs_totalnas
# All NA-heavy ones W1 so no W1 analysis done.
# 3 BL (1st, 21st and 37th), 1 ER (40th obs) missing

plot(df_fast_wide$Timepoint,df_fast_wide$Triglycerides, xlab="Timepoint", ylab="Triglycerides")
```

# Datasets: BL, BL $\to$ EF, EF $\to$ ER

```{r, include=FALSE}
# Baseline data
bl_data=df_fast_wide[Timepoint=="BL",]
rownames(bl_data)=bl_data$ID
bl_data=bl_data[,-c(1,2)]
df_baseline_data=dplyr::select(bl_data,-matches("Weight|Waist|SBP|DBP|^Hb$|Ht|Neutr|Sodium|Potassium|Chloride|Calcium|ALT|^HDL$"))
(NAs_cols=which(colMeans(is.na(df_baseline_data))!=0)) # Bililrubin and UrineSG
colSums(is.na(df_baseline_data[,NAs_cols]))
df_baseline_1=df_baseline_data[,-NAs_cols]
rm(bl_data)

# BL to EF fractionalised-percentage changes 
# (Note: think about logs for skews ???)
blef_change=1+((df_fast_wide[Timepoint=="EF",-c(1,2)]-df_fast_wide[Timepoint=="BL",-c(1,2)])/df_fast_wide[Timepoint=="BL",-c(1,2)])
blef_change_data=dplyr::select(blef_change,-matches("Weight|Waist|SBP|DBP|^Hb$|Ht|Neutr|Sodium|Potassium|Chloride|Calcium|ALT|^HDL$"))
rownames(blef_change_data)=df_fast_wide[rownames(blef_change_data),"ID"]
# View(blef_change_data) #Uncomment to see table
# hist(blef_change_data[,"Weight"]) #Uncomment to see for skews/log-changes
(NAs_cols=which(colMeans(is.na(blef_change_data))!=0)) # Bililrubin and UrineSG
colSums(is.na(blef_change_data[,NAs_cols]))
df_blef_1=blef_change_data[,-NAs_cols]
rm(blef_change_data,blef_change)

# EF to ER fractionalised-percentage changes 
# (Note: think about logs for skews ???)
efer_change=1+((df_fast_wide[Timepoint=="ER",-c(1,2)]-df_fast_wide[Timepoint=="EF",-c(1,2)])/df_fast_wide[Timepoint=="EF",-c(1,2)])
efer_change_data=dplyr::select(efer_change,-matches("Weight|Waist|SBP|DBP|^Hb$|Ht|Neutr|Sodium|Potassium|Chloride|Calcium|ALT|^HDL$"))
rownames(efer_change_data)=df_fast_wide[rownames(efer_change_data),"ID"]
# View(efer_change_data) #Uncomment to see table
# hist(efer_change_data[,"Weight"]) #Uncomment to see for skews/log-changes
(NAs_cols=which(colMeans(is.na(efer_change_data))!=0)) # 1 ER - Bilirubin only
df_efer_1 = efer_change_data[,-NAs_cols]
rm(efer_change_data,efer_change)

detach(df_fast_wide)
```

## Mean-Centering datasets
```{r}
# Mean-centering: https://link.springer.com/article/10.3758/s13428-015-0624-x
# Helps interecept interpretation, preserving slope and also helps a bit with multicollinearity
centr.df_blef_1=as.data.frame(scale(df_blef_1,center=T,scale=F))
centr.df_efer_1=as.data.frame(scale(df_efer_1,center=T,scale=F))
centr.df_baseline_1=as.data.frame(scale(df_baseline_1, center=T, scale=F))
```


# Opposite direction deltas

## Glucose
```{r}
##Delta GLUCOSE##

## Only BLEF
grps.blef=func_delta_frac_subgroups(df_blef_1,Glucose)
incr_grp=grps.blef$increased_grp; n_inc=grps.blef$num_increased_grp
decr_grp=grps.blef$decreased_grp; n_dec=grps.blef$num_decreased_grp

# Increased group
# Decreased group
## With baseline
gluc_bl = cbind(df_blef_1$Glucose, df_baseline_1)
colnames(gluc_bl)[1]="Glucose.BLEF"

grps.bio_bl.blef=func_delta_frac_subgroups(gluc_bl, Glucose.BLEF)
incr_grp=grps.bio_bl.blef$increased_grp; n_inc=grps.bio_bl.blef$num_increased_grp
decr_grp=grps.bio_bl.blef$decreased_grp; n_dec=grps.bio_bl.blef$num_decreased_grp

# Increased group
# Decreased group
func_cor_scatter_plot(decr_grp, c("Triglycerides","SHBG", "Trig_HDL_Ratio"), 
                 title="Data: BL, Grp: Gluc_BLEF Dec")
```

## Insulin
```{r}
##Delta INSULIN##

## Only BLEF
grps.blef=func_delta_frac_subgroups(df_blef_1,Insulin)
incr_grp=grps.blef$increased_grp; n_inc=grps.blef$num_increased_grp
decr_grp=grps.blef$decreased_grp; n_dec=grps.blef$num_decreased_grp

# Increased group

func_cor_scatter_plot(incr_grp, c("BMI","IGFBP.2", "IGF.1"), 
                 title="Data: BLEF, Grp: Ins Inc")
# Decreased group
func_cor_scatter_plot(decr_grp, c("Glucose","Beta.hydroxy.butyrate"), 
                 title="Data: BLEF, Grp: Ins Dec")
## With baseline
ins_bl = cbind(df_blef_1$Insulin, df_baseline_1)
colnames(ins_bl)[1]="Insulin.BLEF"

grps.bio_bl.blef=func_delta_frac_subgroups(ins_bl, Insulin.BLEF)
incr_grp=grps.bio_bl.blef$increased_grp; n_inc=grps.bio_bl.blef$num_increased_grp
decr_grp=grps.bio_bl.blef$decreased_grp; n_dec=grps.bio_bl.blef$num_decreased_grp

# Increased group
func_cor_scatter_plot(incr_grp, c("Insulin.BLEF","IGF.1"), 
                 title="Data: BL, Grp: Ins_BLEF Inc")

# Decreased group
func_cor_scatter_plot(decr_grp, c("FGF.21", "BMI"), 
                 title="Data: BL, Grp: Ins_BLEF Dec")
```

## Triglycerides
```{r}
##Delta TG##

## Only BLEF
grps.blef=func_delta_frac_subgroups(df_blef_1,Triglycerides)
incr_grp=grps.blef$increased_grp; n_inc=grps.blef$num_increased_grp
decr_grp=grps.blef$decreased_grp; n_dec=grps.blef$num_decreased_grp

# Increased group
func_cor_scatter_plot(incr_grp, c("Beta.hydroxy.butyrate", "Glucose"),
                 title="Data: BLEF, Grp: Tg Inc")


# Decreased group
func_cor_scatter_plot(decr_grp, c("Beta.hydroxy.butyrate", "Cholesterol"),
                 title="Data: BLEF, Grp: Tg Dec")

## With baseline
tg_bl = cbind(df_blef_1$Triglycerides, df_baseline_1)
colnames(tg_bl)[1]="Tg.BLEF"

grps.bio_bl.blef=func_delta_frac_subgroups(tg_bl, Tg.BLEF)
incr_grp=grps.bio_bl.blef$increased_grp; n_inc=grps.bio_bl.blef$num_increased_grp
decr_grp=grps.bio_bl.blef$decreased_grp; n_dec=grps.bio_bl.blef$num_decreased_grp


# Increased group
# Decreased group
```

## LDL
```{r}
##Delta LDL##

## Only BLEF
grps.blef=func_delta_frac_subgroups(df_blef_1,LDL)
incr_grp=grps.blef$increased_grp; n_inc=grps.blef$num_increased_grp
decr_grp=grps.blef$decreased_grp; n_dec=grps.blef$num_decreased_grp

# Increased group
func_cor_scatter_plot(incr_grp, c("Beta.hydroxy.butyrate", "Glucose"),
                 title="Data: BLEF, Grp: LDL Inc (Part I)")

func_cor_scatter_plot(incr_grp, c("FGF.21","AlkPhos."),
                 title="Data: BLEF, Grp: LDL Inc (Part II)")

# Decreased group

## With baseline
ldl_bl = cbind(df_blef_1$LDL, df_baseline_1)
colnames(ldl_bl)[1]="LDL.BLEF"

grps.bio_bl.blef=func_delta_frac_subgroups(ldl_bl, LDL.BLEF)
incr_grp=grps.bio_bl.blef$increased_grp; n_inc=grps.bio_bl.blef$num_increased_grp
decr_grp=grps.bio_bl.blef$decreased_grp; n_dec=grps.bio_bl.blef$num_decreased_grp


# Increased group

func_cor_scatter_plot(incr_grp, c("IGFBP.2","CO2"),
                 title="Data: BL, Grp: LDL_BLEF Inc (Part I)")

func_cor_scatter_plot(incr_grp, c("Triglycerides", "SHBG","Trig_HDL_Ratio"),
                 title="Data: BL, Grp: LDL_BLEF Inc (Part II)")


# Decreased group
func_cor_scatter_plot(decr_grp, c("LDL", "SHBG"),
                 title="Data: BL, Grp: LDL_BLEF Dec (Part II)")

func_cor_scatter_plot(decr_grp, c("BMI","FGF.21","AST"),
                 title="Data: BL, Grp: LDL_BLEF Dec (Part III)")
```