---
title: "PF_CSV with PhenoAge"
author: ''
date: "SJ"
output:
  html_document:
    toc: yes
    number_sections: true
    toc_depth: 5
  pdf_document:
    toc: yes
    number_sections: true
    toc_depth: 5
urlcolor: blue
linkcolor: red
---
```{r global_options, include=FALSE}
rm(list = ls()) ##AM reset workspace - good practice to run clean
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment=NA, warning=FALSE, message = FALSE,results = 'hide')
knitr::opts_knit$set(root.dir = "C:/Users/SJ/Desktop/Backup Mac/Desktop/STUDY/Usyd Statistician")
```

# Source functions

```{r}
source("./Functions/Biomarker_functions_SJ.R")
```

# Working Directory & Datasets
```{r, message=FALSE}
setwd("./Andrius Prolonged Fasting")
##AM please suppress trash outputs with chunk option: message=FALSE
##AM sometimes you need this to suppress stubborn outputs
#```{r results='hide', fig.keep='high'}

df_fast_wide = read.csv("Blood_changes_across_timepoints_rows_wide.csv")

# df_fast_wide = read.csv("../DATA/Blood_changes_across_timepoints_rows_wide_AM.csv",header=T,sep=",")
# df_fast_frac_wide = read.csv("../DATA/Blood_changes_across_timepoints_rows_wide_fractions_AM.csv",header=T,sep=",")

# drop rows containing W1 timepoint - we will not use it
# df_fast_wide <- subset(df_fast_wide, Timepoint != "W1")
# df_fast_frac_wide <- subset(df_fast_frac_wide, Timepoint != "W1.f")

# working with df_fast_wide data
attach(df_fast_wide)
df_fast_wide$ID = as.factor(ID)
df_fast_wide$Timepoint=as.factor(Timepoint)
```

# PhenoAge

```{r}
### With PhenoAge
phenoAge=func_PhenotypicAge(df_fast_wide)
df_fast_wide=cbind(df_fast_wide,phenoAge)
write.csv(df_fast_wide,"./Andrius Prolonged Fasting/Blood_changes_across_timepoints_rows_wide_SJ.csv",row.names = T)
detach(df_fast_wide)
```