---
title: "PF_DFs"
author: 'Non-imputed'
date: "SJ"
output:
  html_document:
    toc: yes
    number_sections: true
    toc_depth: 5
  pdf_document:
    toc: yes
    number_sections: true
    toc_depth: 5
urlcolor: blue
linkcolor: red
---
```{r global_options, include=FALSE}
rm(list = ls()) ##AM reset workspace - good practice to run clean
library(dplyr)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, warning=FALSE)
knitr::opts_knit$set(root.dir = "G:/My Drive/Prolonged_Fasting_Study")
theme_update(plot.title = element_text(hjust = 0.5)) # center plots
```


# Working Directory & Datasets
```{r, message=FALSE}
setwd("./DATA")
##AM please suppress trash outputs with chunk option: message=FALSE
##AM sometimes you need this to suppress stubborn outputs
#```{r results='hide', fig.keep='high'}

df_fast_wide = read.csv("Blood_changes_across_timepoints_rows_wide_AM.csv")
df_fast_frac_wide = read.csv("Blood_changes_across_timepoints_rows_fractions_AM.csv")

# df_fast_wide = read.csv("../DATA/Blood_changes_across_timepoints_rows_wide_AM.csv",header=T,sep=",")
# df_fast_frac_wide = read.csv("../DATA/Blood_changes_across_timepoints_rows_wide_fractions_AM.csv",header=T,sep=",")

# drop rows containing W1 timepoint - we will not use it
# df_fast_wide <- subset(df_fast_wide, Timepoint != "W1")
# df_fast_frac_wide <- subset(df_fast_frac_wide, Timepoint != "W1.f")

# working with df_fast_wide data
attach(df_fast_wide)
df_fast_wide$ID = as.factor(ID)
df_fast_wide$Timepoint=as.factor(Timepoint)
```

## Missing values?

```{r}
##AM add print() arguments to label each output
(total_missing=sum(is.na(df_fast_wide)))
(NAs_observations=which(rowMeans(is.na(df_fast_wide))!=0))
rowMeans(is.na(df_fast_wide[NAs_observations,])) # out of 50 observations
obs_totalnas=rowSums(is.na(df_fast_wide[NAs_observations,]))
names(obs_totalnas)=df_fast_wide[NAs_observations,"Timepoint"];obs_totalnas
# All NA-heavy ones W1 so no W1 analysis done.
# 3 BL (1st, 21st and 37th), 1 ER (40th obs) missing

plot(df_fast_wide$Timepoint,df_fast_wide$Triglycerides, xlab="Timepoint", ylab="Triglycerides")
```

# Datasets: BL, BL $\to$ EF, EF $\to$ ER

```{r, include=FALSE}
# Baseline data
bl_data=df_fast_wide[Timepoint=="BL",]
rownames(bl_data)=bl_data$ID
bl_data=bl_data[,-c(1,2)]
df_baseline_data=dplyr::select(bl_data,-matches("Weight|Waist|SBP|DBP|^Hb$|Ht|Neutr|Sodium|Potassium|Chloride|Calcium|ALT"))
(NAs_cols=which(colMeans(is.na(df_baseline_data))!=0)) # Bililrubin and UrineSG
colSums(is.na(df_baseline_data[,NAs_cols]))
df_baseline_1=df_baseline_data[,-NAs_cols]
rm(bl_data,df_baseline_data)

# BL to EF fractionalised-percentage changes 
# (Note: think about logs for skews ???)
blef_change=df_fast_wide[Timepoint=="EF",-c(1,2)]/df_fast_wide[Timepoint=="BL",-c(1,2)]
blef_change_data=dplyr::select(blef_change,-matches("Weight|Waist|SBP|DBP|^Hb$|Ht|Neutr|Sodium|Potassium|Chloride|Calcium|ALT"))
rownames(blef_change_data)=df_fast_wide[rownames(blef_change_data),"ID"]
# View(blef_change_data) #Uncomment to see table
# hist(blef_change_data[,"Weight"]) #Uncomment to see for skews/log-changes
(NAs_cols=which(colMeans(is.na(blef_change_data))!=0)) # Bililrubin and UrineSG
colSums(is.na(blef_change_data[,NAs_cols]))
df_blef_1=blef_change_data[,-NAs_cols]
rm(blef_change_data,blef_change)

# EF to ER fractionalised-percentage changes 
# (Note: think about logs for skews ???)
efer_change=df_fast_wide[Timepoint=="ER",-c(1,2)]/df_fast_wide[Timepoint=="EF",-c(1,2)]
efer_change_data=dplyr::select(efer_change,-matches("Weight|Waist|SBP|DBP|^Hb$|Ht|Neutr|Sodium|Potassium|Chloride|Calcium|ALT"))
rownames(efer_change_data)=df_fast_wide[rownames(efer_change_data),"ID"]
# View(efer_change_data) #Uncomment to see table
# hist(efer_change_data[,"Weight"]) #Uncomment to see for skews/log-changes
(NAs_cols=which(colMeans(is.na(efer_change_data))!=0)) # 1 ER - Bilirubin only
df_efer_1 = efer_change_data[,-NAs_cols]
rm(efer_change_data,efer_change)

detach(df_fast_wide)
```

## Mean-Centering datasets
```{r, include=FALSE}
# Mean-centering: https://link.springer.com/article/10.3758/s13428-015-0624-x
# Helps interecept interpretation, preserving slope and also helps a bit with multicollinearity
centr.df_blef_1=as.data.frame(scale(df_blef_1,center=T,scale=F))
centr.df_efer_1=as.data.frame(scale(df_efer_1,center=T,scale=F))
centr.df_baseline_1=as.data.frame(scale(df_baseline_1, center=T, scale=F))
```

```{r}
setwd("./SCRIPTS")
save(df_fast_wide, df_blef_1, df_baseline_1, df_efer_1, 
     centr.df_baseline_1, centr.df_blef_1, centr.df_blef_1, centr.df_efer_1,
     file="RDA_Non_imputed_dfs_SJ.rda")
print("End of code.")
knitr::knit_exit()
```